---
title: "Core Summaries"
author: "Lisa Zorn"
date: "Friday, August 01, 2014"
output:
  html_document:
    css: custom.css
    toc: yes
# runtime: shiny
---

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

# Initialization: Set the workspace and load needed libraries
```{r Initialization}
library(knitr)
library(ggplot2)
library(scales)
library(xtable)
library(dplyr)
library(reshape2)
opts_knit$set(root.dir="C:/Users/lzorn/Documents/2010_04_ZZZ")
# if running outside of knitr
setwd("C:/Users/lzorn/Documents/2010_04_ZZZ")
```

# Household files

## Read the household files and land use file.

There are two household files:

 * the model input file from the synthesized household/population (http://analytics.mtc.ca.gov/foswiki/Main/PopSynHousehold)
 * the model output file (http://analytics.mtc.ca.gov/foswiki/Main/Household)

The land use file: http://analytics.mtc.ca.gov/foswiki/Main/TazData

```{r ReadHouseholds}
input.pop.households <- read.table(file = "hhFile.p2011s3a1.2010.csv", header=TRUE, sep=",")
input.ct.households  <- read.table(file = "householdData_3.csv", header=TRUE, sep = ",")
walk_subzone_labels  <- c("Cannot walk to transit (more than two-thirds of a mile away)",
                          "Short-walk to transit (less than one-third of a mile)",
                          "Long-walk to transit (between one-third and two-thirds of a mile)")

input.tazData <- read.table(file="tazData.csv", header=TRUE, sep=",")
county_labels <- c("San\nFrancisco","San\nMateo","Santa\nClara","Alameda",
                   "Contra\nCosta","Solano","Napa","Sonoma","Marin")
```

## Join them

Rename/drop some columns and join them on household id. Also join with tazData to get the super district and county.

```{r JoinHouseholds}
input.pop.households <- select(input.pop.households, HHID, PERSONS, hworkers, huniv, hpresch, hschpred, hschdriv)
input.ct.households <- select(input.ct.households, -jtf_choice)
input.tazData <- select(input.tazData, ZONE, SD, COUNTY)

# rename
names(input.pop.households)[names(input.pop.households)=="HHID"] <- "hh_id"
names(input.tazData)[names(input.tazData)=="ZONE"] <- "taz"

households <- inner_join(input.pop.households, input.ct.households)
households <- inner_join(households, input.tazData)
# wrap as a d data frame tbl so it's nicer for printing
households <- tbl_df(households)
```

## Check that autos is always set
```{r CheckAutosSet}
sum(is.na(households$autos))
stopifnot(sum(is.na(households$autos))==0)
```

## Recode a few new variables and summarize
```{r RecodeHouseholdVars}
# incQ are Income Quartiles
households    <- mutate(households, incQ=1*(income<30000) + 
                                         2*((income>=30000)&(income<60000)) +
                                         3*((income>=60000)&(income<100000)) +
                                         4*(income>=100000))

# workers are hworkers capped at 4
households    <- mutate(households, workers=4*(hworkers>=4) + hworkers*(hworkers<4))
worker_labels <- c("Zero", "One", "Two", "Three", "Four or more")

# kidsNoDr is 1 if the household has children in the househole that don't drive
# (either pre-school age or school age)
households    <- mutate(households, kidsNoDr=(hpresch>=1)|(hschpred>=1))
```

# Trip Files

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/IndividualTrip

```{r ReadTrips}
trips           <- read.table(file="indivTripData_3.csv", header=TRUE, sep=",")
joint_trips     <- read.table(file="jointTripData_3.csv", header=TRUE, sep=",")

trips           <- select(input.ct.trips, hh_id, person_id, orig_taz, dest_taz,
                          trip_mode, orig_purpose, dest_purpose, depart_hour)
trips           <- mutate(trips, 
                          amode=1*(trip_mode==7) +                                           # walk
                                2*(trip_mode==8) +                                           # bike
                                3*((trip_mode>=9)&(trip_mode<=13)) +                         # walk trn walk
                                4*((trip_mode>=14)&(trip_mode<=18)&(orig_purpose=='Home')) + # drive trn walk
                                5*((trip_mode>=14)&(trip_mode<=18)&(dest_purpose=='Home')),  # walk trn drive
                          wlk_trip=1*(amode==1),
                          bik_trip=1*(amode==2),
                          wtr_trip=1*(amode==4),
                          dtr_trip=1*(amode==3) + 1*(amode==5),
                          timeCode=1*(depart_hour<6) +                                       # EA
                                   2*((depart_hour>5)&(depart_hour<10)) +                    # AM
                                   3*((depart_hour>9)&(depart_hour<15)) +                    # MD
                                   4*((depart_hour>14)&(depart_hour<19)) +                   # PM
                                   5*((depart_hour>18)),                                     # EV
                          active=0
                          )
colnames(trips)
```

```{r attachActive}
timeCode_labels <- c("EA","AM","MD","PM","EV")
timeCode_values <- c(1,2,3,4,5)
trips2          <- trips
for (timeCode_val in timeCode_values) {
  timeCode    <- timeCode_labels[timeCode_val]
  activeSkims <- read.table(file = paste0("ActiveTimeSkimsDatabase",timeCode,".csv"), header=TRUE, sep=",")
  # rename columns for join
  names(activeSkims)[names(activeSkims)=="orig"] <- "orig_taz"  
  names(activeSkims)[names(activeSkims)=="dest"] <- "dest_taz"
  activeSkims <- mutate(activeSkims, timeCode=timeCode_val)
  
  # left join trips to the skims
  trips2 <- left_join(trips2, activeSkims, by=c("orig_taz","dest_taz","timeCode"))
  # assign active value if we can
  trips2 <- mutate(trips2,
                   active2=((timeCode==timeCode_val)*(amode==1)*walk) +
                           ((timeCode==timeCode_val)*(amode==2)*bike) +
                           ((timeCode==timeCode_val)*(amode==3)*wTrnW) +
                           ((timeCode==timeCode_val)*(amode==4)*dTrnW) +
                           ((timeCode==timeCode_val)*(amode==5)*wTrnD)
                   )
  sum(is.na(trips2$active2))
  trips2$active2[is.na(trips2$active2)] <- 0
  trips2 <- mutate(trips2, active=active+active2)
  sum(trips2$active==0)
  trips2 <- select(trips2, -walk, -bike, -wTrnW, -dTrnW, -wTrnD)
}
```

# Auto Ownership Summaries

## Auto Ownership Summary By County
```{r AutoOwnershipSummaryByCountyGraph, fig.width=8, fig.height=5}
# summarize
autoown_summary <- summarise(group_by(households, SD, COUNTY, autos, incQ, walk_subzone, workers, kidsNoDr), freq=n())
auto_labels <- c("Zero automobiles", "One automobile", "Two automobiles",
                 "Three automobiles", "Four or more automobiles")

# by County
autoown_summary_county <- summarise(group_by(autoown_summary, COUNTY, autos), freq=sum(freq))
ggplot(autoown_summary_county, aes(x=as.factor(COUNTY), y=freq, fill=as.factor(autos))) +
  geom_bar(stat="identity", position="fill") +
  scale_y_continuous(labels=percent, breaks=seq(0,1,0.1)) +
  scale_x_discrete(labels=county_labels) +
  xlab("County of Residence") + 
  ylab("Share of Households") +
  ggtitle("Households by Automobile Ownership and County of Residence") +
  scale_fill_discrete(name="", 
                      breaks=c(4,3,2,1,0), 
                      labels=rev(auto_labels))
```

```{r AutoOwnershipSummaryByCountyTable, results='asis', echo=FALSE}
aos_ac_table                     <- dcast(autoown_summary_county, autos ~ COUNTY, value.var="freq")
# make rownames readable
stopifnot(aos_ac_table$autos == c(0,1,2,3,4))
rownames(aos_ac_table)           <- auto_labels
aos_ac_table                     <- select(aos_ac_table, -autos)
# add Sum row and name it Everyone
aos_ac_table_summed              <- addmargins(as.table(as.matrix(aos_ac_table)), 1, FUN=sum)
rownames(aos_ac_table_summed)[6] <- "Everyone"
# name columns
colnames(aos_ac_table_summed)    <- county_labels
aos_ac_xtable                    <- xtable(apply(aos_ac_table_summed, 2, prettyNum, big.mark=","), 
                                         align="lrrrrrrrrr",
                                         caption="Automobile Ownership Levels by County")
print(aos_ac_xtable, type="html")


# shares
aos_ac_table_share              <- prop.table(as.table(as.matrix(aos_ac_table)),2)
aos_ac_table_share              <- addmargins(aos_ac_table_share, 1, FUN=sum)
rownames(aos_ac_table_share)[6] <- "Everyone"
colnames(aos_ac_table_share)    <- county_labels
aos_ac_xtable_share             <- xtable(apply(aos_ac_table_share, 2, function(u) sprintf("%.1f%%",u*100)),
                                        align="lrrrrrrrrr",
                                        caption="Automobile Ownership Share by County")
rownames(aos_ac_xtable_share)   <- append(auto_labels,"Everyone")
print(aos_ac_xtable_share, type="html")
```

## Auto Ownership Summary By Workers
```{r AutoOwnershipSummaryByWorkersGraph, fig.width=8, fig.height=5}
# by Workers
autoown_summary_workers <- summarise(group_by(autoown_summary, workers, autos), freq=sum(freq))

ggplot(autoown_summary_workers, aes(x=as.factor(workers), y=freq, fill=as.factor(autos))) +
  geom_bar(stat="identity", position="fill") +
  scale_y_continuous(labels=percent, breaks=seq(0,1,0.1)) +
  scale_x_discrete(labels=worker_labels) +
  xlab("Workers in the Household") + 
  ylab("Share of Households") +
  ggtitle("Households by Automobile Ownership and Number of Workers") +
  scale_fill_discrete(name="", 
                      breaks=c(4,3,2,1,0), 
                      labels=rev(auto_labels))
```

```{r AutoOwnershipSummaryByWorkersTable, results='asis', echo=FALSE}
aos_aw_table <- dcast(autoown_summary_workers, autos ~ workers, value.var="freq")
# make rownames readable
stopifnot(aos_aw_table$autos == c(0,1,2,3,4))
rownames(aos_aw_table)           <- auto_labels
aos_aw_table                     <- select(aos_aw_table, -autos)
# add Sum row and name it Everyone
aos_aw_table_summed              <- addmargins(as.table(as.matrix(aos_aw_table)), 1, FUN=sum)
rownames(aos_aw_table_summed)[6] <- "Everyone"
# name columns
colnames(aos_aw_table_summed)    <- worker_labels
aosw_xtable                      <- xtable(apply(aos_aw_table_summed, 2, prettyNum, big.mark=","), 
                                           align="lrrrrr",
                                           caption="Automobile Ownership Levels by Number of Workers")
print(aosw_xtable, type="html")


# shares
aos_aw_table_share              <- prop.table(as.table(as.matrix(aos_aw_table)),2)
aos_aw_table_share              <- addmargins(aos_aw_table_share, 1, FUN=sum)
rownames(aos_aw_table_share)[6] <- "Everyone"
colnames(aos_aw_table_share)    <- worker_labels
aosw_xtable_share               <- xtable(apply(aos_aw_table_share, 2, function(u) sprintf("%.1f%%",u*100)),
                                          align="lrrrrr",
                                          caption="Automobile Ownership Share by Number of Workers")
rownames(aosw_xtable_share)     <- append(auto_labels,"Everyone")
print(aosw_xtable_share, type="html")
```

## Average Auto Ownership Summary By County and Transit Availability
```{r AutoOwnershipSummaryByCountyTrnAvailGraph, fig.width=8, fig.height=5}
# by Workers
autoown_summary_trncoutny <- summarise(group_by(autoown_summary, COUNTY, walk_subzone, autos), freq=sum(freq))
autoown_summary_trncoutny <- mutate(autoown_summary_trncoutny, totautos=autos*freq)
# ggplot(autoown_summary_workers, aes(x=as.factor(workers), y=freq, fill=as.factor(autos))) +
#  geom_bar(stat="identity", position="fill") +
#  scale_y_continuous(labels=percent, breaks=seq(0,1,0.1)) +
#  scale_x_discrete(labels=worker_labels) +
#  xlab("Workers in the Household") + 
#  ylab("Share of Households") +
#  ggtitle("Households by Automobile Ownership and Number of Workers") +
#  scale_fill_discrete(name="", 
#                      breaks=c(4,3,2,1,0), 
#                      labels=rev(auto_labels))
```

```{r AutoOwnershipSummaryByCountyTrnAvailTable, results='asis', echo=FALSE}
# households in walk_subzones x county
aos_wc_table                  <- dcast(autoown_summary_trncoutny, walk_subzone ~ COUNTY, 
                                       value.var="freq", fun.aggregate=sum)
stopifnot(aos_wc_table$walk_subzone == c(0,1,2))
rownames(aos_wc_table)        <- walk_subzone_labels
aos_wc_table                  <- select(aos_wc_table, -walk_subzone)
aos_wc_table_summed           <- addmargins(as.table(as.matrix(aos_wc_table)), 1, FUN=sum)

# total autos in walk_subzones x county
aos_wc_table_totaut           <- dcast(autoown_summary_trncoutny, walk_subzone ~ COUNTY, 
                                       value.var="totautos", fun.aggregate=sum)
stopifnot(aos_wc_table_totaut$walk_subzone == c(0,1,2))
rownames(aos_wc_table_totaut) <- walk_subzone_labels
aos_wc_table_totaut           <- select(aos_wc_table_totaut, -walk_subzone)
aos_wc_table_totaut_summed    <- addmargins(as.table(as.matrix(aos_wc_table_totaut)), 1, FUN=sum)

# avg autos in walk_subzones x county
aos_wc_table_avgaut           <- aos_wc_table_totaut_summed/aos_wc_table_summed
rownames(aos_wc_table_avgaut)[4] <- "Everyone"
colnames(aos_wc_table_avgaut) <- county_labels
aos_wc_xtable_avgaut          <- xtable(apply(aos_wc_table_avgaut, 2, function(u) sprintf("%.2f",u)), 
                                        align="lrrrrrrrrr",
                                        caption=paste("Average Automobile Ownership Levels",
                                                      "(capped at 4 per household) by County ",
                                                      "and Walk-to-Transit Accessibility"))
rownames(aos_wc_xtable_avgaut)<- append(walk_subzone_labels, "Everyone")
print(aos_wc_xtable_avgaut, type="html")

# number of households in walk_subzones x county
rownames(aos_wc_table_summed)[4] <- "Everyone"
colnames(aos_wc_table_summed) <- county_labels
aos_wc_xtable_summed          <- xtable(apply(aos_wc_table_summed, 2, prettyNum, big.mark=","), 
                                        align="lrrrrrrrrr",
                                        caption=paste("Number of households by County",
                                                      "and Walk-to-Transit Accessibility"))
rownames(aos_wc_xtable_summed)<- append(walk_subzone_labels, "Everyone")
print(aos_wc_xtable_summed, type="html")

```