---
title: "Transit System Stats"
author: "David Ory"
output: 
  html_document:
    theme: cosmo
    toc: yes
---


## Administration

#### Purpose
Summarizes transit system -- separately for the system and segmented by operator and technology -- statistics for use with the state of good repair assessments across travel model simulations. 

#### Outputs
1.  A CSV database with logical names of system-wide statistics.
2.  A CSV database with logical names of operator-technology-specific statistics.

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
suppressMessages(library(dplyr))
library(stringr)
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote I/O
```{r remote-io}
TARGET_DIR_ARRAY <- c("M:/Application/Model One/RTP2013/Scenarios/2000_03_YYY/OUTPUT/trn/",
                      "M:/Application/Model One/RTP2013/Scenarios/2005_03_YYY/OUTPUT/trn/",
                      "M:/Application/Model One/RTP2013/Scenarios/2010_03_YYY/OUTPUT/trn/",
                      "M:/Application/Model One/RTP2013/Scenarios/Round 05 -- Final/2015_03_116/OUTPUT/trn/",
                      "M:/Application/Model One/RTP2013/Scenarios/Round 05 -- Final/2040_03_116/OUTPUT/trn/"
                      )

F_INPUT_MODE <- "transit-modes.csv"

OUTPUT_DIR <- "M:/Application/Model One/RTP2013/Scenarios/Round 05 -- Across Scenarios/bespoke/"
F_OUTPUT_SYSTEM = paste(OUTPUT_DIR, "transit-system-stats.csv", sep = "")
F_OUTPUT_MODE   = paste(OUTPUT_DIR, "transit-mode-operator-stats.csv", sep = "")
```

#### Parameters
```{r parameters}
YEAR_ARRAY <- c(2000, 2005, 2010, 2015, 2040)
SCENARIO_ID_ARRAY <- c("2000_03_YYY", "2005_03_YYY", "2010_03_YYY", "2015_03_116", "2040_03_116")

time_period = c("ea", "am", "md", "pm", "ev")
hours = c(3, 4, 5, 4, 8)
hours_in_periods_df <- data.frame(time_period, hours)
hours_in_periods_df <- hours_in_periods_df %>%
  mutate(time_period = paste(time_period))

```

#### System-wide calculator method
```{r calculator-method}
Compute_Transit_System_Statistics <- function(target_dir, year_int, scenario_id) {
  
  # read in the file
  input_file_name <- paste(target_dir, "trnline.csv", sep = "")
  input_df <- read.table(file = input_file_name, header = TRUE, sep = ",", stringsAsFactors = FALSE)
  
  # passenger miles
  passenger_miles <- sum(input_df$passenger.miles)
  
  # vehicle_miles
  supply_df <- input_df %>%
    filter(path.id == "ea_wlk_com_wlk" | 
             path.id == "am_wlk_com_wlk" | 
             path.id == "md_wlk_com_wlk" | 
             path.id == "pm_wlk_com_wlk" | 
             path.id == "ev_wlk_com_wlk") %>%
    mutate(route_speed = line.dist / line.time * 60) %>%
    mutate(route_vehicles = ceiling(line.time / frequency)) %>%
    mutate(route_vehicle_miles_per_hour = route_vehicles * route_speed) %>%
    mutate(time_period = str_sub(path.id, start = 1L, end = 2L))
  
  supply_df <- left_join(supply_df, hours_in_periods_df, by = c("time_period"))
  
  supply_df <- supply_df %>%
    mutate(route_vehicle_miles = route_vehicle_miles_per_hour * hours)
  
  revenue_vehicle_miles <- sum(supply_df$route_vehicle_miles)
  
  # passenger miles weighted headway
  passenger_miles_weighted_headway <- weighted.mean(input_df$passenger.miles, input_df$frequency)
  
  # passenger boardings
  boardings <- sum(input_df$total.boardings)
  
  # route length
  average_route_length <- mean(supply_df$line.dist)
  
  # build and return a dataframe
  return_df <- data.frame(year = c(year_int), 
                          scenario_id = c(scenario_id), 
                          passenger_miles = c(passenger_miles),
                          revenue_vehicle_miles = c(revenue_vehicle_miles),
                          passenger_miles_weighted_headway = c(passenger_miles_weighted_headway),
                          boardings = c(boardings),
                          average_unweighted_route_length = c(average_route_length))
  
  return(return_df)
  
  }


```

#### Mode/operator-specific calculator method
```{r mode-calculator-method}
Compute_Transit_Mode_Operator_Statistics <- function(target_dir, year_int, scenario_id) {
  
  # read in the relevant file
  input_file_name <- paste(target_dir, "trnline.csv", sep = "")
  input_df <- read.table(file = input_file_name, header = TRUE, sep = ",", stringsAsFactors = FALSE)
  
  # fix mode 37 AND 85 bugS (both AC Transit)
  input_df <- input_df %>%
    mutate(mode = ifelse(mode == 37, 30, mode)) %>%
    mutate(mode = ifelse(mode == 85, 30, mode))
  
  # passenger miles
  activity_df <- input_df %>%
    rename(passenger_miles = passenger.miles, boardings = total.boardings)
  
  # vehicle_miles
  supply_df <- input_df %>%
    filter(path.id == "ea_wlk_com_wlk" | 
             path.id == "am_wlk_com_wlk" | 
             path.id == "md_wlk_com_wlk" | 
             path.id == "pm_wlk_com_wlk" | 
             path.id == "ev_wlk_com_wlk") %>%
    mutate(route_speed = line.dist / line.time * 60) %>%
    mutate(route_vehicles = ceiling(line.time / frequency)) %>%
    mutate(route_vehicle_miles_per_hour = route_vehicles * route_speed) %>%
    mutate(time_period = str_sub(path.id, start = 1L, end = 2L))
  
  supply_df <- left_join(supply_df, hours_in_periods_df, by = c("time_period"))
  
  supply_df <- supply_df %>%
    mutate(route_vehicle_miles = route_vehicle_miles_per_hour * hours)
  
  # passenger miles weighted headway
  activity_df <- activity_df %>%
    mutate(pax_miles_times_freq = passenger_miles * frequency)
  
  activity_sum <- activity_df %>%
    group_by(mode) %>%
    summarise(passenger_miles = sum(passenger_miles),
              pax_miles_times_freq_sum = sum(pax_miles_times_freq),
              boardings = sum(boardings))
  
  activity_sum <- activity_sum %>%
    mutate(passenger_miles_weighted_headway = pax_miles_times_freq_sum / passenger_miles) %>%
    select(-pax_miles_times_freq_sum)
  
  supply_sum <- supply_df %>%
    group_by(mode) %>%
    summarise(revenue_vehicle_miles = sum(route_vehicle_miles),
              average_unweighted_route_length = mean(line.dist))
  
  sum_df <- merge(activity_sum, supply_sum, all = TRUE) 
  
  sum_df <- sum_df %>%
    ungroup() %>%
    mutate(year = year_int) %>%
    mutate(scenario_id = scenario_id)
  
  } 

```

#### Apply the system-wide method
```{r apply-system-method}
system_df <- Compute_Transit_System_Statistics(TARGET_DIR_ARRAY[1], YEAR_ARRAY[1], SCENARIO_ID_ARRAY[1])
for (i in 2:length(TARGET_DIR_ARRAY)){
  
  target_dir <- TARGET_DIR_ARRAY[i]
  year_int <- YEAR_ARRAY[i]
  scenario_id <- SCENARIO_ID_ARRAY[i]
  
  df <- Compute_Transit_System_Statistics(target_dir, year_int, scenario_id)
  system_df <- rbind(system_df, df)
  
  }

```

#### Apply the mode/operator method
```{r apply-mode-operator-method}
mode_oper_df <- Compute_Transit_Mode_Operator_Statistics(TARGET_DIR_ARRAY[1], YEAR_ARRAY[1], SCENARIO_ID_ARRAY[1])
for (i in 2:length(TARGET_DIR_ARRAY)){
  
  target_dir <- TARGET_DIR_ARRAY[i]
  year_int <- YEAR_ARRAY[i]
  scenario_id <- SCENARIO_ID_ARRAY[i]
  
  df <- Compute_Transit_Mode_Operator_Statistics(target_dir, year_int, scenario_id)
  mode_oper_df <- rbind(mode_oper_df, df)
  
  }

# join with mode codes
mode_df <- read.table(file = F_INPUT_MODE, header = TRUE, sep = ",", stringsAsFactors = FALSE)
mode_oper_df <- left_join(mode_oper_df, mode_df, by = c("mode"))

```

#### Data writes
```{r data-writes}
write.csv(system_df, file = F_OUTPUT_SYSTEM, row.names = FALSE, quote = F)
write.csv(mode_oper_df, file = F_OUTPUT_MODE, row.names = FALSE, quote = F)
```

