---
title: "VMT by income and mode"
author: "Lisa Zorn"
output: 
  html_document:
    theme: cosmo
    toc: yes
---

## Administration

#### Purpose
Prepares a bespoke summary of travel model output.  Specifically, summarizes all VMT (including the drive component of drive to transit,
which requires reading ITHIM skims rather than just relying on simple skims for the transit trips) by income and mode.

### Outputs
1. A CSV database3 with logical names.

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
suppressMessages(library(dplyr))
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Paramaters
```{r parameters}
SAMPLING_RATE = 0.50
```

#### Remote file locations
```{r remote-dir}
MODEL_DIR   <- "D:/Projects/2040_06_695"
TARGET_DIR  <- MODEL_DIR
OUTPUT_DIR  <- "M:/Application/Model One/RTP2017/Scenarios/2040_06_695/OUTPUT/bespoke"
SCENARIO_NAME = "2040_06_695"
```

### Load, join, compute
```{r data-manips}
# Load person trips
load(file.path(MODEL_DIR, "updated_output","trips.rdata"))

# We only need some columns: trip_mode, orig_taz, dest_taz, timeCode, incQ, incQ_label, distance, num_participants
trips <- select(trips,
                trip_mode, orig_taz, dest_taz, timeCode, incQ, incQ_label, distance, num_participants, orig_purpose)

# We only want those with VMT - so drive or drive to transit
auto_trips <- subset(trips, trip_mode<=6)
dtrn_trips <- subset(trips, trip_mode>=14)

# Sum VMT for auto trips by splitting joint person miles traveled into vehicle miles traveled
auto_trips <- mutate(auto_trips,
                     vmt_indiv=(num_participants==1)*(trip_mode==1)*distance +
                               (num_participants==1)*(trip_mode==2)*distance +
                               (num_participants==1)*(trip_mode==3)*(distance/2.0) +
                               (num_participants==1)*(trip_mode==4)*(distance/2.0) +
                               (num_participants==1)*(trip_mode==5)*(distance/3.25) +
                               (num_participants==1)*(trip_mode==6)*(distance/3.25),
                     vmt_joint=(num_participants>1)*(distance/num_participants),
                     vmt=vmt_indiv+vmt_joint,
                     vehicle_trips = (num_participants==1)*(trip_mode==1)*1.0 +
                                     (num_participants==1)*(trip_mode==2)*1.0 +
                                     (num_participants==1)*(trip_mode==3)*(1.0/2.0) +
                                     (num_participants==1)*(trip_mode==4)*(1.0/2.0) +
                                     (num_participants==1)*(trip_mode==5)*(1.0/3.25) +
                                     (num_participants==1)*(trip_mode==6)*(1.0/3.25) +
                                     (num_participants>1)*(1.0/num_participants))
auto_vmt <- summarise(group_by(auto_trips, incQ, incQ_label, trip_mode),
                      person_trips  = n(),
                      vehicle_trips = sum(vehicle_trips),
                      vmt           = sum(vmt))


# For transit, we need to join to ITHIM skims since they have drive distances
# This is from utilities\PBA40\ITHIM\PerCapitaDailyTravelDistanceTime.R
add_ithim_skims <- function(this_timeperiod, input_trips) {
  # separate the relevant and irrelevant tours/trips
  relevant   <- input_trips %>% filter(timeCode == this_timeperiod)
  irrelevant <- input_trips %>% filter(timeCode != this_timeperiod)
  
  # read in the relevant skim
  skim_file   <- file.path(TARGET_DIR,"database",paste0("IthimSkimsDatabase",this_timeperiod,".csv"))
  ithim_skims <- read.table(file=skim_file, header=TRUE, sep=",")
  # standardize column names
  ithim_skims <- ithim_skims %>% rename(orig_taz=orig, dest_taz=dest)
  
  # Left join tours to the skims
  relevant <- left_join(relevant, ithim_skims, by=c("orig_taz","dest_taz"))
  # assign values if we can
  relevant <- relevant %>% 
    mutate(ivtB=(trn_mode=='wTrnW')*ivtB_wTrnW +
                (trn_mode=='dTrnW')*ivtB_dTrnW +
                (trn_mode=='wTrnD')*ivtB_wTrnD) %>%
    mutate(ivtR=(trn_mode=='wTrnW')*ivtR_wTrnW +
                (trn_mode=='dTrnW')*ivtR_dTrnW +
                (trn_mode=='wTrnD')*ivtR_wTrnD) %>%
    mutate(distB=(trn_mode=='wTrnW')*distB_wTrnW +
                 (trn_mode=='dTrnW')*distB_dTrnW +
                 (trn_mode=='wTrnD')*distB_wTrnD) %>%
    mutate(distR=(trn_mode=='wTrnW')*distR_wTrnW +
                 (trn_mode=='dTrnW')*distR_dTrnW +
                 (trn_mode=='wTrnD')*distR_wTrnD) %>%
    mutate(ddist=(trn_mode=='dTrnW')*ddist_dTrnW +
                 (trn_mode=='wTrnD')*ddist_wTrnD) %>%
    mutate(dtime=(trn_mode=='dTrnW')*dtime_dTrnW +
                 (trn_mode=='wTrnD')*dtime_wTrnD)
  
  # re-code missing as zero
  relevant <- relevant %>%
    mutate(ivtB = ifelse(ivtB < -990, 0, ivtB)) %>%
    mutate(ivtR = ifelse(ivtR < -990, 0, ivtR)) %>%
    mutate(distB = ifelse(distB < -990, 0, distB)) %>%
    mutate(distR = ifelse(distR < -990, 0, distR)) %>%
    mutate(ddist = ifelse(ddist < -990, 0, distB)) %>%
    mutate(dtime = ifelse(dtime < -990, 0, distR)) %>%
    mutate(ivt_trn  = ivtB+ivtR) %>%
    mutate(dist_trn = distB + distR)

  print(paste("For", 
              this_timeperiod, 
              "assigned", 
              prettyNum(sum(!is.na(relevant$ivt_trn)),big.mark=","),
              "ivts and ",
              prettyNum(sum(!is.na(relevant$dist_trn)),big.mark=","),
              "dists"))
  
  print(paste("  -> Total zero ivts:",
              prettyNum(sum(relevant$ivt_trn==0),big.mark=",")))
  print(paste("  -> Total zero dists:",
              prettyNum(sum(relevant$dist_trn==0),big.mark=",")))
  # clean-up
  relevant <- relevant %>% select(-walk_wTrnW,-ivtB_wTrnW,-ivtR_wTrnW,-wait_wTrnW,-distB_wTrnW,-distR_wTrnW,
                                  -walk_wTrnD,-ivtB_wTrnD,-ivtR_wTrnD,-wait_wTrnD,-distB_wTrnD,-distR_wTrnD,-dtime_wTrnD,-ddist_wTrnD,
                                  -walk_dTrnW,-ivtB_dTrnW,-ivtR_dTrnW,-wait_dTrnW,-distB_dTrnW,-distR_dTrnW,-dtime_dTrnW,-ddist_dTrnW)
  
  return_list <- rbind(relevant, irrelevant)
  return(return_list)
}

# Some trips are drive-transit-walk and some are walk-transit-drive depending on the origin purpose
dtrn_trips <- mutate(dtrn_trips, trn_mode = ifelse(orig_purpose=='Home', 'dTrnW','wTrnD')) %>%
              mutate(vehicle_trips = (num_participants==1)*1.0 +
                                     (num_participants>1)*(1.0/num_participants))

# add the drive distance from the ITHIM skims
dtrn_trips <- mutate(dtrn_trips, ivtB=0, ivtR=0, distB=0, distR=0, ddist=0, dtime=0, ivt_trn=0, dist_trn=0)
dtrn_trips <- add_ithim_skims('EA', dtrn_trips)
dtrn_trips <- add_ithim_skims('AM', dtrn_trips)
dtrn_trips <- add_ithim_skims('MD', dtrn_trips)
dtrn_trips <- add_ithim_skims('PM', dtrn_trips)
dtrn_trips <- add_ithim_skims('EV', dtrn_trips)
dtrn_trips <- mutate(dtrn_trips, vmt=vehicle_trips*ddist)

dtrn_vmt <- summarise(group_by(dtrn_trips, incQ, incQ_label, trip_mode),
                      person_trips  = n(),
                      vehicle_trips = sum(vehicle_trips),
                      vmt           = sum(vmt))

all_vmt <- rbind(auto_vmt, dtrn_vmt)
# sampling rate affects all
all_vmt <- all_vmt %>%
  mutate(person_trips  = person_trips / SAMPLING_RATE,
         vehicle_trips = vehicle_trips / SAMPLING_RATE,
         vmt           = vmt / SAMPLING_RATE)
```

#### Write to disk
```{r write-disk}
F_OUTPUT = file.path(OUTPUT_DIR, "vmt-income-mode.csv")
write.csv(all_vmt, file = F_OUTPUT, row.names = FALSE, quote = F)
```

