---
title: "TripsSimple"
author: "Lisa Zorn"
date: "Wednesday, November 05, 2014"
output: html_document
---

This html is output by https://github.com/MetropolitanTransportationCommission/Travel-Model-One-Utilities/blob/master/CoreSummaries/DataRequests/Trips_origSD_destSD_departhour_tripmode.Rmd

Simple script to summarize *trips* by
 * Superdistrict of origin
 * Superdistrict of Destination
 * Departure Hour
 * Travel mode
 
Note: Much of this code is from CoreSummaries.Rmd but I'm running it in
directories without CoreSummaries output so it's simpler just to make a new one.

It should be easy to run via `runDataRequestSummary.bat`.

Outputs file into subdir `OUTPUT` if it exists, otherwise into a `core_summaries`.

# Initialization: Set the workspace and load needed libraries
```{r Initialization, results='hold', error=FALSE}
library(knitr)
library(ggplot2)
library(scales)
library(dplyr)
library(reshape2)

# For RStudio, these can be set in the .Rprofile
TARGET_DIR   <- Sys.getenv("TARGET_DIR")  # The location of the input files
ITER         <- Sys.getenv("ITER")        # The iteration of model outputs to read

stopifnot(nchar(TARGET_DIR  )>0)
stopifnot(nchar(ITER        )>0)

cat("TARGET_DIR  = ",TARGET_DIR)
cat("ITER        = ",ITER)
```

# Land Use

The land use file: http://analytics.mtc.ca.gov/foswiki/Main/TazData
```{r readLandUse, error=FALSE}
file_rundir       <- file.path(TARGET_DIR,"landuse","tazData.csv")
file_archive      <- file.path(TARGET_DIR,"OUTPUT","tazData.csv")
if (file.exists(file_rundir)) {
  tazData         <- read.table(file=file_rundir, header=TRUE, sep=",")
} else {
  tazData         <- read.table(file=file_archive, header=TRUE, sep=",")
}
tazData           <- select(tazData, ZONE, SD, COUNTY, PRKCST, OPRKCST)
names(tazData)[names(tazData)=="ZONE"] <- "taz"
```

# Trips

## Read Individual Trips and recode a few variables

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/IndividualTrip

```{r ReadIndividualTrips, error=FALSE}
file_rundir       <- file.path(TARGET_DIR,"main",paste0("indivTripData_",ITER,".csv"))
file_archive      <- file.path(TARGET_DIR,"OUTPUT",paste0("indivTripData_",ITER,".csv"))
if (file.exists(file_rundir)) {
  indiv_trips     <- tbl_df(read.table(file=file_rundir, header=TRUE, sep=","))
} else {
  indiv_trips     <- tbl_df(read.table(file=file_archive, header=TRUE, sep=","))
}
indiv_trips     <- select(indiv_trips, orig_taz, dest_taz, trip_mode, depart_hour)
indiv_trips     <- mutate(indiv_trips, num_participants=1)

print(paste("Read",prettyNum(nrow(indiv_trips),big.mark=",")))
```

## Read Joint Trips and recode a few variables

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/JointTrip

```{r ReadJointTrips, error=FALSE}
file_rundir       <- file.path(TARGET_DIR,"main",paste0("jointTripData_",ITER,".csv"))
file_archive      <- file.path(TARGET_DIR,"OUTPUT",paste0("jointTripData_",ITER,".csv"))
if (file.exists(file_rundir)) {
  joint_trips     <- tbl_df(read.table(file=file_rundir, header=TRUE, sep=","))
} else {
  joint_trips     <- tbl_df(read.table(file=file_archive, header=TRUE, sep=","))
}
joint_trips     <- select(joint_trips, orig_taz, dest_taz, trip_mode,
                          num_participants, depart_hour)

print(paste("Read",prettyNum(nrow(joint_trips),big.mark=","),
            "joint trips or ",prettyNum(sum(joint_trips$num_participants),big.mark=","),
            "joint person trips"))
```

## Append them and add the superdistricts, and save

```{r CombineTrips, error=FALSE}
trips <- tbl_df(rbind(indiv_trips, joint_trips))
remove(indiv_trips,joint_trips)

trips <- left_join(trips,
                  mutate(tazData, orig_taz=taz, orig_SD=SD, orig_COUNTY=COUNTY) %.% 
                  select(orig_taz, orig_SD, orig_COUNTY),
                  by=c("orig_taz"))

trips <- left_join(trips,
                  mutate(tazData, dest_taz=taz, dest_SD=SD, dest_COUNTY=COUNTY) %.% 
                  select(dest_taz, dest_SD, dest_COUNTY),
                  by=c("dest_taz"))

trips_summary <- summarise(group_by(trips,
                                    orig_SD, orig_COUNTY,
                                    dest_SD, dest_COUNTY,
                                    trip_mode, depart_hour), 
                            freq     = n(),
                            num_participants = sum(num_participants))

if (file.exists(file.path(TARGET_DIR,"OUTPUT"))) {
  write.table(trips_summary, file.path(TARGET_DIR,"OUTPUT","Trips_origSD_destSD_departhour_tripmode.csv"), sep=",", row.names=FALSE)
} else {
  # write results in TARGET_DIR/core_summaries
  if (!file.exists(file.path(TARGET_DIR,"core_summaries"))) {
    dir.create(file.path(TARGET_DIR,"core_summaries"))
  }
  write.table(trips_summary, file.path(TARGET_DIR,"core_summaries","Trips_origSD_destSD_departhour_tripmode.csv"), sep=",", row.names=FALSE)
}
```