---
title: "Core Summaries"
author: "Lisa Zorn"
date: "Friday, August 01, 2014"
output:
  html_document:
    css: custom.css
    toc: yes
# runtime: shiny
---

# Initialization: Set the workspace and load needed libraries
```{r Initialization, results='hold', error=FALSE}
library(knitr)
library(ggplot2)
library(scales)
library(xtable)
library(dplyr)
library(reshape2)

# For RStudio, these can be set in the .Rprofile
TARGET_DIR   <- Sys.getenv("TARGET_DIR")  # The location of the input files
POPSYN_HH    <- Sys.getenv("POPSYN_HH")   # The synthesized households file
POPSYN_PERS  <- Sys.getenv("POPSYN_PERS") # The synthesized persons file
ITER         <- Sys.getenv("ITER")        # The iteration of model outputs to read

TARGET_DIR   <- gsub("\\\\","/",TARGET_DIR) # switch slashes around

# run test mode?  It's a short version
TEST        <- FALSE
if (TEST) {
  TARGET_DIR  <- "C:/Users/lzorn/Documents/2020_03_116"
  POPSYN_HH   <- "hhFile.p2011s6g.2020"
  POPSYN_PERS <- "personFile.p2011s6g.2020"
  ITER        <- "1"
}

stopifnot(nchar(TARGET_DIR  )>0)
stopifnot(nchar(POPSYN_HH   )>0)
stopifnot(nchar(POPSYN_PERS )>0)
stopifnot(nchar(ITER        )>0)

# write results in TARGET_DIR/core_summaries
if (!file.exists(file.path(TARGET_DIR,"core_summaries"))) {
  dir.create(file.path(TARGET_DIR,"core_summaries"))
}
# write tables in TARGET_DIR/updated_output
if (!file.exists(file.path(TARGET_DIR,"updated_output"))) {
  dir.create(file.path(TARGET_DIR,"updated_output"))
}

cat("TARGET_DIR  = ",TARGET_DIR)
cat("POPSYN_HH   = ",POPSYN_HH)
cat("POPSYN_PERS = ",POPSYN_PERS)
cat("ITER        = ",ITER)
```

# Lookups

For time periods, see http://analytics.mtc.ca.gov/foswiki/Main/TimePeriods
For counties, see http://analytics.mtc.ca.gov/foswiki/Main/TazData
For walk_subzones, see http://analytics.mtc.ca.gov/foswiki/Main/Household

```{r timeCodeDefinitions, error=FALSE}
######### time periods 
LOOKUP_TIMEPERIOD    <- data.frame(timeCodeNum=c(1,2,3,4,5),
                                   timeperiod_label=c("Early AM","AM Peak","Midday","PM Peak","Evening"),
                                   timeperiod_abbrev=c("EA","AM","MD","PM","EV"))
# no factors -- joins don't work
LOOKUP_TIMEPERIOD$timeCodeNum       <- as.integer(LOOKUP_TIMEPERIOD$timeCodeNum)
LOOKUP_TIMEPERIOD$timeperiod_label  <- as.character(LOOKUP_TIMEPERIOD$timeperiod_label)
LOOKUP_TIMEPERIOD$timeperiod_abbrev <- as.character(LOOKUP_TIMEPERIOD$timeperiod_abbrev)

######### counties
LOOKUP_COUNTY        <- data.frame(COUNTY=c(1,2,3,4,5,6,7,8,9),
                                   county_name=c("San Francisco","San Mateo","Santa Clara","Alameda",
                                                 "Contra Costa","Solano","Napa","Sonoma","Marin"))
LOOKUP_COUNTY$COUNTY      <- as.integer(LOOKUP_COUNTY$COUNTY)
LOOKUP_COUNTY$county_name <- as.character(LOOKUP_COUNTY$county_name)

######### walk subzones
LOOKUP_WALK_SUBZONE  <- data.frame(walk_subzone=c(0,1,2),
                                   walk_subzone_label=c("Cannot walk to transit (more than two-thirds of a mile away)",
                                                        "Short-walk to transit (less than one-third of a mile)",
                                                        "Long-walk to transit (between one-third and two-thirds of a mile)"))
LOOKUP_WALK_SUBZONE$walk_subzone <- as.integer(LOOKUP_WALK_SUBZONE$walk_subzone)

######### person types
LOOKUP_PTYPE         <- data.frame(ptype=c(1,2,3,4,5,6,7,8),
                                   ptype_label=c("Full-time worker","Part-time worker","College student","Non-working adult",
                                                 "Retired","Driving-age student","Non-driving-age student",
                                                 "Child too young for school"))
LOOKUP_PTYPE$ptype   <- as.integer(LOOKUP_PTYPE$ptype)

```

# Land Use

The land use file: http://analytics.mtc.ca.gov/foswiki/Main/TazData
```{r readLandUse, error=FALSE}
tazData           <- read.table(file=file.path(TARGET_DIR,"landuse","tazData.csv"), header=TRUE, sep=",")
tazData           <- select(tazData, ZONE, SD, COUNTY, PRKCST, OPRKCST)
tazData           <- left_join(tazData, LOOKUP_COUNTY, by=c("COUNTY"))
names(tazData)[names(tazData)=="ZONE"] <- "taz"

```

# Household files

## Read the household files and land use file

There are two household files:

 * the model input file from the synthesized household/population (http://analytics.mtc.ca.gov/foswiki/Main/PopSynHousehold)
 * the model output file (http://analytics.mtc.ca.gov/foswiki/Main/Household)

```{r ReadHouseholds, error=FALSE}
input.pop.households <- read.table(file = file.path(TARGET_DIR,"popsyn",paste0(POPSYN_HH,".csv")), 
                                   header=TRUE, sep=",")
input.ct.households  <- read.table(file = file.path(TARGET_DIR,"popsyn",paste0("householdData_",ITER,".csv")), 
                                   header=TRUE, sep = ",")

```

## Join them

Rename/drop some columns and join them on household id. Also join with tazData to get the super district and county.

```{r JoinHouseholds, error=FALSE}
input.pop.households <- select(input.pop.households, HHID, PERSONS, hworkers, huniv, hpresch, 
                               hschpred, hschdriv)
input.ct.households  <- select(input.ct.households, -jtf_choice)

# rename
names(input.pop.households)[names(input.pop.households)=="HHID"] <- "hh_id"

households <- inner_join(input.pop.households, input.ct.households, "hh_id")
households <- inner_join(households, tazData, "taz")
# wrap as a d data frame tbl so it's nicer for printing
households <- tbl_df(households)
# clean up
remove(input.pop.households, input.ct.households)
```


## Recode a few new variables

Create the following new household variables:
  * income quartiles (`incQ`)
  * worker categories (`workers`)
  * dummy for households with children that don't drive (`kidsNoDr`)
  * auto sufficiency (`autoSuff`)
  * walk subzone label (`walk_subzone_label`)
  
```{r RecodeHouseholdVars, error=FALSE}
# incQ are Income Quartiles
LOOKUP_INCQ          <- data.frame(incQ=c(1,2,3,4),
                                   incQ_label=c("Less than $30k","$30k to $60k","$60k to $100k","More than $100k"))
LOOKUP_INCQ$incQ     <- as.integer(LOOKUP_INCQ$incQ)

households    <- mutate(households, incQ=1*(income<30000) + 
                                         2*((income>=30000)&(income<60000)) +
                                         3*((income>=60000)&(income<100000)) +
                                         4*(income>=100000))
households    <- left_join(households, LOOKUP_INCQ, by=c("incQ"))

# workers are hworkers capped at 4
households    <- mutate(households, workers=4*(hworkers>=4) + hworkers*(hworkers<4))
WORKER_LABELS <- c("Zero", "One", "Two", "Three", "Four or more")

# kidsNoDr is 1 if the household has children in the househole that don't drive
# (either pre-school age or school age)
households    <- mutate(households, kidsNoDr=(hpresch>=1)|(hschpred>=1))

# auto sufficiency
LOOKUP_AUTOSUFF          <- data.frame(autoSuff=c(0,1,2),
                                autoSuff_label=c("Zero automobiles","Automobiles < workers","Automobiles >= workers"))
LOOKUP_AUTOSUFF$autoSuff <- as.integer(LOOKUP_AUTOSUFF$autoSuff)

households    <- mutate(households, autoSuff=1*((autos>0)&(autos<hworkers)) +
                                             2*((autos>0)&(autos>=hworkers)))
households    <- left_join(households, LOOKUP_AUTOSUFF, by=c("autoSuff"))

# walk subzone label
households    <- left_join(households, LOOKUP_WALK_SUBZONE, by=c("walk_subzone"))
```

# Person files

There are two person files:

 * the model input file from the synthesized household/population (http://analytics.mtc.ca.gov/foswiki/Main/PopSynPerson)
 * the model output file (http://analytics.mtc.ca.gov/foswiki/Main/Person)

## Read the person files
```{r ReadPersons, error=FALSE}
input.pop.persons    <- read.table(file = file.path(TARGET_DIR,"main",paste0(POPSYN_PERS,".csv")), 
                                   header=TRUE, sep=",")
input.ct.persons     <- read.table(file = file.path(TARGET_DIR,"main",paste0("personData_",ITER,".csv")),
                                   header=TRUE, sep = ",")
```

## Join them
```{r JoinPersons, error=FALSE}
# Rename
names(input.pop.persons)[names(input.pop.persons)=="HHID"] <- "hh_id"
names(input.pop.persons)[names(input.pop.persons)=="PERID"] <- "person_id"

# Inner join so persons must be present in both.  So only simulated persons stay.
persons              <- inner_join(input.pop.persons, input.ct.persons, by=c("hh_id", "person_id"))
# Get incQ from Households
persons              <- left_join(persons, select(households, hh_id, incQ, incQ_label), by=c("hh_id"))

# Person type label
persons              <- left_join(persons, LOOKUP_PTYPE, by=c("ptype"))

# wrap as a d data frame tbl so it's nicer for printing
persons              <- tbl_df(persons)
# clean up
remove(input.pop.persons, input.ct.persons)
```

# Tours

## Read Individual Tours

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/IndividualTour

```{r readIndividualTours, error=FALSE}
indiv_tours     <- tbl_df(read.table(file=file.path(TARGET_DIR,"main",
                                                    paste0("indivTourData_",ITER,".csv")), 
                                     header=TRUE, sep=","))

# Add income from household table
indiv_tours     <- left_join(indiv_tours, select(households, hh_id, income, incQ, incQ_label), by=c("hh_id"))
indiv_tours     <- mutate(indiv_tours, num_participants=1)

# Add in County, Superdistrict, Parking Cost from TAZ Data for the tour destination
indiv_tours     <- left_join(indiv_tours,
                             mutate(tazData, dest_taz=taz, dest_COUNTY=COUNTY, dest_county_name=county_name, 
                                    dest_SD=SD, PRKCST, OPRKCST) %.% 
                               select(dest_taz, dest_COUNTY, dest_county_name, dest_SD, PRKCST, OPRKCST),
                             by=c("dest_taz"))

# Add free-parking choice from persons table
indiv_tours   <- left_join(indiv_tours, select(persons, person_id, fp_choice), by=c("person_id"))

# Compute the tour parking rate
indiv_tours   <- mutate(indiv_tours, parking_rate=ifelse(tour_category=='MANDATORY',PRKCST,OPRKCST))

# Free parking for work tours if fp_choice==1
indiv_tours   <- mutate(indiv_tours, parking_rate=ifelse((substr(tour_purpose,0,4)=='work')*(fp_choice==1),0.0,parking_rate))
```

## Read Joint Tours

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/JointTour

```{r readJointTours, error=FALSE}
joint_tours    <- tbl_df(read.table(file=file.path(TARGET_DIR,"main",
                                                   paste0("jointTourData_",ITER,".csv")), 
                                    header=TRUE, sep=","))
# Add Income from household table
joint_tours    <- left_join(joint_tours, select(households, hh_id, income, incQ, incQ_label), by=c("hh_id"))

# Add in County, Superdistrict, Parking Cost from TAZ Data for the tour destination
joint_tours     <- left_join(joint_tours,
                             mutate(tazData, dest_taz=taz, dest_COUNTY=COUNTY, dest_county_name=county_name, 
                                    dest_SD=SD, PRKCST, OPRKCST) %.% 
                               select(dest_taz, dest_COUNTY, dest_county_name, dest_SD, PRKCST, OPRKCST),
                             by=c("dest_taz"))

# Add parking rate
joint_tours    <- mutate(joint_tours, parking_rate=OPRKCST)
```

## Function to add Cost to Tours or Trips

Function `add_cost`: attaches the cost skims for the given time period abbreviation `this_timeperiod`
to the given tours data.frame (note: this can be the trips data frame). Joining is done on the columns 
`orig_taz`, `dest_taz`, and `timeCode`.

The function fills in values to the column `cost` and `cost_fail` for the given time
period based on the column `costMode`.

Pass `reverse_od` as TRUE to do return trip; this uses the dest taz as the origin and the origin
taz as the dest, and it uses wTrnD cost rather than dTrnW.

Tallies failed costs (expected cost but -999 found) in the field cost_fail

```{r addCostColumn}
add_cost <- function(this_timeperiod, tours, reverse_od=FALSE) {
  
  costSkims   <- read.table(file = file.path(TARGET_DIR,"database",
                                             paste0("CostSkimsDatabase",
                                                    this_timeperiod,".csv")), 
                            header=TRUE, sep=",")
  # rename columns for join -- reversing if needed (return trip)
  if (reverse_od) {
    names(costSkims)[names(costSkims)=="orig"] <- "dest_taz"  
    names(costSkims)[names(costSkims)=="dest"] <- "orig_taz"    
  } else {
    names(costSkims)[names(costSkims)=="orig"] <- "orig_taz"  
    names(costSkims)[names(costSkims)=="dest"] <- "dest_taz"
  }
  # don't recode -999 as 0 -- we'll deal with it after join
  costSkims   <- mutate(costSkims,
                        timeCode = this_timeperiod)
  
  # Left join tours to the skims
  tours <- left_join(tours, costSkims, by=c("orig_taz","dest_taz","timeCode"))
  # assign active value if we can to new column active2
  tours <- mutate(tours,
                  cost2=((timeCode==this_timeperiod)*(costMode==1)*da    ) +
                        ((timeCode==this_timeperiod)*(costMode==2)*daToll) +
                        ((timeCode==this_timeperiod)*(costMode==3)*s2    ) +
                        ((timeCode==this_timeperiod)*(costMode==4)*s2Toll) +
                        ((timeCode==this_timeperiod)*(costMode==5)*s3    ) +
                        ((timeCode==this_timeperiod)*(costMode==6)*s3Toll) +
                        ((timeCode==this_timeperiod)*(costMode==7)*0.0   ) +
                        ((timeCode==this_timeperiod)*(costMode==8)*0.0   ) +
                        ((timeCode==this_timeperiod)*(costMode==9)*wTrnW ) +
                        ((timeCode==this_timeperiod)*(costMode==10)*(1-reverse_od)*dTrnW) +
                        ((timeCode==this_timeperiod)*(costMode==10)*(reverse_od   *wTrnD)) +
                        ((timeCode==this_timeperiod)*(costMode==11)*wTrnD)
                  )

  tours <- mutate(tours,
                  cost_fail2 = ifelse(cost2==-999,1,0),    # tally the cost_fail
                  cost2      = ifelse(cost2==-999,0,cost2) # reset the -999 to zero
                  )
      
  print(paste("For", this_timeperiod, "assigned", prettyNum(sum(!is.na(tours$cost2)),big.mark=","),
              "costs, with", prettyNum(sum(!is.na(tours$cost2)&(tours$cost2>0)),big.mark=","),
              "nonzero values"))
  tours$cost2[is.na(tours$cost2)]           <- 0           # make the unassigned ones 0 for summing
  tours$cost_fail2[is.na(tours$cost_fail2)] <- 0           # make the unassigned ones 0 for summing
  tours <- mutate(tours, cost=cost+cost2,        # sum
                         cost_fail=cost_fail+cost_fail2)
  print(paste("  -> total nonzero costs:",prettyNum(sum(tours$cost!=0),big.mark=",")))
  tours <- select(tours, -da, -daToll, -s2, -s2Toll, -s3, -s3Toll, -wTrnW, -dTrnW, -wTrnD, -cost2, -cost_fail2) # rid new variables
}
```

## Function to add Distance to Tours or Trips

Function `add_distance`: attaches the distance skims the given time period abbreviation `this_timeperiod`
to the given tours data frame (note: this can be the trips data frame). Joining is done on the columns 
`orig_taz`, `dest_taz`, and `timeCode`. The function fills in values to the column `distance` for the given
time period based on the column `distance_mode`.

```{r addDistanceColumn}
add_distance <- function(this_timeperiod, tours) {
  
  distSkims   <- read.table(file = file.path(TARGET_DIR,"database",
                                             paste0("DistanceSkimsDatabase",
                                                    this_timeperiod,".csv")), 
                            header=TRUE, sep=",")
  # rename columns for join
  names(distSkims)[names(distSkims)=="orig"] <- "orig_taz"  
  names(distSkims)[names(distSkims)=="dest"] <- "dest_taz"
  
  distSkims   <- mutate(distSkims,
                        timeCode = this_timeperiod,
                        da       = ifelse(da    == -999,0,da    ),
                        daToll   = ifelse(daToll== -999,0,daToll),
                        s2       = ifelse(s2    == -999,0,s2    ),
                        s2Toll   = ifelse(s2Toll== -999,0,s2Toll),
                        s3       = ifelse(s3    == -999,0,s3    ),
                        s3Toll   = ifelse(s3Toll== -999,0,s3Toll),
                        walk     = ifelse(walk  == -999,0,walk  ),
                        bike     = ifelse(bike  == -999,0,bike ))
  
  # Left join tours to the skims
  tours <- left_join(tours, distSkims, by=c("orig_taz","dest_taz","timeCode"))
  
  # assign active value if we can to new column active2
  tours <- mutate(tours,
                  distance2=((timeCode==this_timeperiod)*(distance_mode==1)*da    ) +
                            ((timeCode==this_timeperiod)*(distance_mode==2)*daToll) +
                            ((timeCode==this_timeperiod)*(distance_mode==3)*s2    ) +
                            ((timeCode==this_timeperiod)*(distance_mode==4)*s2Toll) +
                            ((timeCode==this_timeperiod)*(distance_mode==5)*s3    ) +
                            ((timeCode==this_timeperiod)*(distance_mode==6)*s3Toll) +
                            ((timeCode==this_timeperiod)*(distance_mode==7)*walk  ) +
                            ((timeCode==this_timeperiod)*(distance_mode==8)*bike  ) +
                            # Use the drive alone time as the distance for transit modes
                            ((timeCode==this_timeperiod)*(distance_mode>=9)*da    )
                  )
  
  print(paste("For", this_timeperiod, "assigned", prettyNum(sum(!is.na(tours$distance2)),big.mark=","),
              "distances, with", prettyNum(sum(!is.na(tours$distance2)&(tours$distance2>0)),big.mark=","),
              "nonzero values"))
  tours$distance2[is.na(tours$distance2)] <- 0         # make the unassigned ones 0 for summing
  tours <- mutate(tours, distance=distance+distance2)  # sum
  print(paste("  -> total nonzero distances:",prettyNum(sum(tours$distance!=0),big.mark=",")))
  tours <- select(tours, -da, -daToll, -s2, -s2Toll, -s3, -s3Toll, -walk, -bike, -distance2) # rid new variables
}
```

## Function to add Time to Tours or Trips

Function `add_time`: attaches the time skims for the given time period `this_timeperiod`
to the given tours data.frame (note: this can be the trips data frame).  Joining is done on the
columns `orig_taz`, `dest_taz`, and `timeCode`.

Fills in values to the column `time` for the given time period based on the column `costMode`.

Pass `reverse_od` as TRUE to do return trip; this uses the dest taz as the origin and the origin
taz as the dest, and it uses wTrnD time rather than dTrnW.

Tallies failed times (expected time but -999 found) in the field `time_fail`.

```{r addTimeColumn}
add_time <- function(this_timeperiod, tours, reverse_od=FALSE) {
  
  timeSkims   <- read.table(file = file.path(TARGET_DIR,"database",
                                             paste0("TimeSkimsDatabase",
                                                    this_timeperiod,".csv")), 
                            header=TRUE, sep=",")
  # rename columns for join -- reversing if needed (return trip)
  if (reverse_od) {
    names(timeSkims)[names(timeSkims)=="orig"] <- "dest_taz"  
    names(timeSkims)[names(timeSkims)=="dest"] <- "orig_taz"    
  } else {
    names(timeSkims)[names(timeSkims)=="orig"] <- "orig_taz"  
    names(timeSkims)[names(timeSkims)=="dest"] <- "dest_taz"
  }
  # don't recode -999 as 0 -- we'll deal with it after join
  timeSkims   <- mutate(timeSkims,
                        timeCode = this_timeperiod)
  
  # Left join tours to the skims
  tours <- left_join(tours, timeSkims, by=c("orig_taz","dest_taz","timeCode"))
  # assign time value if we can to new column time2
  tours <- mutate(tours,
                  time2=((timeCode==this_timeperiod)*(costMode==1)*da    ) +
                        ((timeCode==this_timeperiod)*(costMode==2)*daToll) +
                        ((timeCode==this_timeperiod)*(costMode==3)*s2    ) +
                        ((timeCode==this_timeperiod)*(costMode==4)*s2Toll) +
                        ((timeCode==this_timeperiod)*(costMode==5)*s3    ) +
                        ((timeCode==this_timeperiod)*(costMode==6)*s3Toll) +
                        ((timeCode==this_timeperiod)*(costMode==7)*walk  ) +
                        ((timeCode==this_timeperiod)*(costMode==8)*bike  ) +
                        ((timeCode==this_timeperiod)*(costMode==9)*wTrnW ) +
                        ((timeCode==this_timeperiod)*(costMode==10)*(1-reverse_od)*dTrnW) +
                        ((timeCode==this_timeperiod)*(costMode==10)*(reverse_od   *wTrnD))
                  )
  
  tours <- mutate(tours,
                  time_fail2 = ifelse(time2==-999,1,0),    # tally the time_fail
                  time2      = ifelse(time2==-999,0,time2) # reset the -999 to zero
                  )
      
  print(paste("For", this_timeperiod, "assigned", prettyNum(sum(!is.na(tours$time2)),big.mark=","),
              "times, with", prettyNum(sum(!is.na(tours$time2)&(tours$time2>0)),big.mark=","),
              "nonzero values"))
  tours$time2[is.na(tours$time2)]           <- 0           # make the unassigned ones 0 for summing
  tours$time_fail2[is.na(tours$time_fail2)] <- 0           # make the unassigned ones 0 for summing
  tours <- mutate(tours, time=time+time2,        # sum
                         time_fail=time_fail+time_fail2)
  print(paste("  -> total nonzero times:",prettyNum(sum(tours$time!=0),big.mark=",")))
  tours <- select(tours, -da, -daToll, -s2, -s2Toll, -s3, -s3Toll, -walk, -bike, -wTrnW, -dTrnW, -wTrnD, -time2, -time_fail2) # rid new variables
}
```


## Function to add Active travel time to tours

Function `add_active`: attaches the active skims of the given time period `this_timeperiod`
to the given trip data.frame (joining on the columns `orig_taz`, `dest_taz`, and `timeCode`).

Fills in values to the column `active` for the given time period.

```{r addActiveColumn}
add_active <- function(this_timeperiod, trips) {

  activeSkims <- read.table(file = file.path(TARGET_DIR,"database",
                                             paste0("ActiveTimeSkimsDatabase",
                                                    this_timeperiod,".csv")),
                            header=TRUE, sep=",")
  # rename columns for join
  names(activeSkims)[names(activeSkims)=="orig"] <- "orig_taz"
  names(activeSkims)[names(activeSkims)=="dest"] <- "dest_taz"
  activeSkims <- mutate(activeSkims,
                        timeCode = this_timeperiod,
                        walk2    = (walk!=-999)*walk,
                        bike2    = (bike!=-999)*bike,
                        wTrnW2   = (wTrnW!=-999)*wTrnW,
                        dTrnW2   = (dTrnW!=-999)*dTrnW,
                        wTrnD2   = (wTrnD!=-999)*wTrnD)
  
  # left join trips to the skims
  trips <- left_join(trips, activeSkims, by=c("orig_taz","dest_taz","timeCode"))
  # assign active value if we can to new column active2
  trips <- mutate(trips,
                  active2=((timeCode==this_timeperiod)*(amode==1)*walk2) +
                          ((timeCode==this_timeperiod)*(amode==2)*bike2) +
                          ((timeCode==this_timeperiod)*(amode==3)*wTrnW2) +
                          ((timeCode==this_timeperiod)*(amode==4)*dTrnW2) +
                          ((timeCode==this_timeperiod)*(amode==5)*wTrnD2)
                  )
  print(paste("For", this_timeperiod, "assigned", prettyNum(sum(!is.na(trips$active2)),big.mark=","),
              "active times, with", prettyNum(sum(!is.na(trips$active2)&(trips$active2>0)),big.mark=","),
              "nonzero values"))
  trips$active2[is.na(trips$active2)] <- 0           # make the unassigned ones 0 for summing
  trips <- mutate(trips, active=active+active2)      # sum
  print(paste("  -> total nonzero active times:",prettyNum(sum(trips$active!=0),big.mark=",")))
  trips <- select(trips, -walk, -bike, -wTrnW, -dTrnW, -wTrnD, -walk2, -bike2, -wTrnW2, -dTrnW2, -wTrnD2, -active2) # rid new variables
}
```
# Trips

## Read Individual Trips and recode a few variables

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/IndividualTrip

```{r ReadIndividualTrips, error=FALSE}
indiv_trips     <- read.table(file=file.path(TARGET_DIR,"main",
                                             paste0("indivTripData_",ITER,".csv")), header=TRUE, sep=",")
indiv_trips     <- select(indiv_trips, hh_id, person_id, tour_id, orig_taz, dest_taz,
                          trip_mode, tour_purpose, orig_purpose, dest_purpose, depart_hour)
```

## Read Joint Trips and recode a few variables

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/JointTrip

```{r ReadJointTrips, error=FALSE}
joint_trips     <- tbl_df(read.table(file=file.path(TARGET_DIR,"main",
                                                    paste0("jointTripData_",ITER,".csv")), 
                                     header=TRUE, sep=","))
joint_trips     <- select(joint_trips, hh_id, tour_id, orig_taz, dest_taz, trip_mode,
                          num_participants, tour_purpose, orig_purpose, dest_purpose, depart_hour)

print(paste("Read",prettyNum(nrow(joint_trips),big.mark=","),
            "joint trips or ",prettyNum(sum(joint_trips$num_participants),big.mark=","),
            "joint person trips"))

## Add `num_participants` to joint_tours
joint_tours     <- left_join(joint_tours, 
                             unique(select(joint_trips, hh_id, tour_id, num_participants)),
                             by=c("hh_id","tour_id"))
```

## Combine individual tours and joint tours together

```{r combineTours, error=FALSE}
tours <- rbind(select(indiv_tours, -person_id, -person_num, -person_type, -atWork_freq, -fp_choice),
               select(joint_tours, -tour_composition, -tour_participants))

# done with this -- joint_tours will be used for unwinding joint trips and then released
remove(indiv_tours)

# add residence TAZ info
tours <- left_join(tours, select(households, hh_id, taz, SD, COUNTY, county_name),
                   by=c("hh_id"))

# add simple_purpose
tours <- mutate(tours, simple_purpose='non-work')
tours$simple_purpose[tours$tour_purpose=='work_low'       ] <- 'work'
tours$simple_purpose[tours$tour_purpose=='work_med'       ] <- 'work'
tours$simple_purpose[tours$tour_purpose=='work_high'      ] <- 'work'
tours$simple_purpose[tours$tour_purpose=='work_very high' ] <- 'work'

tours$simple_purpose[tours$tour_purpose=='school_grade'   ] <- 'school'
tours$simple_purpose[tours$tour_purpose=='school_high'    ] <- 'school'

tours$simple_purpose[tours$tour_purpose=='university'     ] <- 'college'

tours$simple_purpose[tours$tour_purpose=='atwork_business'] <- 'at work'
tours$simple_purpose[tours$tour_purpose=='atwork_eat'     ] <- 'at work'
tours$simple_purpose[tours$tour_purpose=='atwork_maint'   ] <- 'at work'

# Calculate the duration
tours   <- mutate(tours, duration=end_hour-start_hour+0.5)

# Parking cost is based on tour duration
tours   <- mutate(tours, parking_cost=parking_rate*duration)
# Distribute costs across shared ride modes (same value used in skims, assignment) for indiv tours
tours   <- mutate(tours, 
                  parking_cost=ifelse((num_participants==1)&((tour_mode==3)|(tour_mode==4)),
                                      parking_cost/1.75,parking_cost))
tours   <- mutate(tours,
                  parking_cost=ifelse((num_participants==1)&((tour_mode==5)|(tour_mode==6)),
                                      parking_cost/2.50,parking_cost))
# Set the transit parking cost to zero
tours   <- mutate(tours, parking_cost=ifelse(tour_mode>6,0.0,parking_cost))
```

## Convert joint trips to joint person trips

Do this by getting the tour participants person ids from the joint_tours table, and unwinding it so that each joint trip
because a trip per partipant.

```{r UnwindJoinTrips, error=FALSE}
# tour participants are person ids separated by spaces -- create a table of hh_id, person_num for them
joint_tour_persons <- data.frame(hh_id=numeric(), tour_id=numeric(), person_num=numeric())
# unwind particpants into table with cols hh_id, tour_id, person_num1, person_num2, ...
participants   <- strsplit(as.character(joint_tours$tour_participants)," ")
max_peeps      <- max(sapply(participants,length))
participants   <- lapply(participants, function(X) c(X,rep(NA, max_peeps-length(X))))
participants   <- data.frame(t(do.call(cbind, participants)))
participants   <- mutate(participants, hh_id=joint_tours$hh_id, tour_id=joint_tours$tour_id)
# melt the persons so they are each on their own row
for (peep in 1:max_peeps) {
  jtp <- melt(participants, id.var=c("hh_id","tour_id"), measure.vars=paste0("X",peep), na.rm=TRUE)
  jtp <- mutate(jtp, person_num=value)
  jtp <- select(jtp, hh_id, tour_id, person_num)
  joint_tour_persons <- rbind(joint_tour_persons, jtp)
}
joint_tour_persons <- transform(joint_tour_persons, person_num=as.numeric(person_num))
# sort by hh_id
joint_tour_persons <- joint_tour_persons[with(joint_tour_persons, order(hh_id, tour_id)),]
# merge with the persons to get the person_id
joint_tour_persons <- left_join(joint_tour_persons, persons, by=c("hh_id","person_num"))

# attach persons to the joint_trips
joint_person_trips <- inner_join(joint_trips, joint_tour_persons, by=c("hh_id", "tour_id"))
# select out person_num and the person table columns
joint_person_trips <- select(joint_person_trips, hh_id, person_id, tour_id, orig_taz, dest_taz, trip_mode,
                             num_participants, tour_purpose, orig_purpose, dest_purpose, depart_hour)
# cleanup
remove(peep,participants,max_peeps,jtp,joint_tours,joint_trips,joint_tour_persons)
```

## Combine Individual Trips and Joint Person Trips
```{r CombineTrips, error=FALSE}
indiv_trips        <- mutate(indiv_trips,        num_participants=1)
trips <- tbl_df(rbind(indiv_trips, joint_person_trips))
print(paste("Combined",prettyNum(nrow(indiv_trips),big.mark=","),
            "individual trips with",prettyNum(nrow(joint_person_trips),big.mark=","),
            "joint trips to make",prettyNum(nrow(trips),big.mark=",")," total trips"))
remove(indiv_trips,joint_person_trips)
```

## Add Variables to Trips

Add some variable to trips, such as:
  * `timeCode`, a recoding of the `depart_hour` for joining skims
  * `home_taz` from household table
  * `incQ` and label from the household table
  * `autoSuff` and label from the household table
  * `walk_subzone` and label from the household table
  * `ptype` and label from persons
  
```{r Trips_AddTimeCode, error=FALSE}
trips <- mutate(trips, 
                timeCodeNum=1*(depart_hour<6) +                                       # EA
                            2*((depart_hour>5)&(depart_hour<10)) +                    # AM
                            3*((depart_hour>9)&(depart_hour<15)) +                    # MD
                            4*((depart_hour>14)&(depart_hour<19)) +                   # PM
                            5*((depart_hour>18))                                      # EV
                )
trips <- left_join(trips, LOOKUP_TIMEPERIOD, by=c("timeCodeNum"))
trips <- select(mutate(trips, timeCode=timeperiod_abbrev), -timeperiod_abbrev)
trips <- left_join(trips,                   
                   mutate(households, home_taz=taz) %.% 
                     select(hh_id, incQ, incQ_label, autoSuff, autoSuff_label,
                            home_taz, walk_subzone, walk_subzone_label),
                   by=c("hh_id"))
trips <- left_join(trips,
                   select(persons, hh_id, person_id, ptype, ptype_label),
                   by=c("hh_id","person_id"))
```

## Add Trip Distance to Trips

Use `add_distance` to add trip distance to trips.
  
```{r Trips_AddDistance, error=FALSE}
trips <- mutate(trips, 
                distance=0.0,
                distance_mode=trip_mode) # use trip_mode for distance

for (timeperiod in LOOKUP_TIMEPERIOD$timeperiod_abbrev) {
  trips <- add_distance(timeperiod, trips)
}
trips <- tbl_df(trips)
trips <- select(trips, -distance_mode)
```

## Add Active Travel time to Trips

```{r Trips_AddActiveColumn, error=FALSE}
# code the Active Mode
trips <- mutate(trips, 
                amode=1*(trip_mode==7) +                                           # walk
                      2*(trip_mode==8) +                                           # bike
                      3*((trip_mode>=9)&(trip_mode<=13)) +                         # walk trn walk
                      4*((trip_mode>=14)&(trip_mode<=18)&(orig_purpose=='Home')) + # drive trn walk
                      5*((trip_mode>=14)&(trip_mode<=18)&(dest_purpose=='Home')),  # walk trn drive
                wlk_trip=1*(amode==1),
                bik_trip=1*(amode==2),
                wtr_trip=1*(amode==3),
                dtr_trip=1*(amode==4) + 1*(amode==5),
                active=0)

# go go gadget: Add active transportation time to trips
for (timeperiod in LOOKUP_TIMEPERIOD$timeperiod_abbrev) {
  trips <- add_active(timeperiod, trips)
}
trips <- tbl_df(trips)
```

## Add Travel Cost and Travel Time to Trips
```{r Trips_AddCostColumn, error=FALSE}
trips <- mutate(trips,
                costMode  =(trip_mode<=8)*trip_mode +         # auto, walk, bike all in the skims
                           # All walk to transit sub modes get best path walk to transit skim
                           (trip_mode>=9)*(trip_mode<=13)*9 +
                           # Drive to transit it from home
                           (trip_mode>=14)*(orig_purpose=='Home')*10 +
                           (trip_mode>=14)*(dest_purpose=='Home')*11,
                cost      =0,
                cost_fail =0,
                time      =0,
                time_fail =0)
for (timeperiod in LOOKUP_TIMEPERIOD$timeperiod_abbrev) {
  trips <- add_cost(timeperiod, trips)
  trips <- add_time(timeperiod, trips)
}
trips <- tbl_df(trips)
```

# Mandatory Locations

The fields are documented here: http://analytics.mtc.ca.gov/foswiki/Main/MandatoryLocation

```{r readMandatoryLocations, error=FALSE}
mandatory_locations <- tbl_df(read.table(file=file.path(TARGET_DIR,"main",
                                                        paste0("wsLocResults_",ITER,".csv")), 
                                         header=TRUE, sep=","))
```

# Summaries

## Active Transportation Summary
```{r ActiveTransportationSummary, error=FALSE}
# we only want some variables - sum them to hh_id/person_id
trips_melt  <- tbl_df(melt(select(trips, hh_id, person_id, wlk_trip, bik_trip, wtr_trip, dtr_trip, active), 
                           id.vars=c("hh_id","person_id")))
trips_dcast <- tbl_df(dcast(trips_melt, hh_id+person_id ~ variable, fun.aggregate=sum))
# left join with persons --> one row per person!
trips_by_person <- left_join(select(persons, hh_id, person_id, ptype, activity_pattern, value_of_time), 
                             trips_dcast, by=c("hh_id","person_id"))

# we're done with these
remove(trips_melt,trips_dcast)

# append autos, taz, COUNTY from households
trips_by_person <- left_join(trips_by_person, select(households, hh_id, autos, taz, county_name), by=c("hh_id"))
# recode zeroAuto
trips_by_person <- mutate(trips_by_person, 
                          zeroAuto=(autos==0),
                          atHomeA=is.na(active))
# set NAs to zero
trips_by_person$wlk_trip[is.na(trips_by_person$wlk_trip)] <- 0
trips_by_person$bik_trip[is.na(trips_by_person$bik_trip)] <- 0
trips_by_person$wtr_trip[is.na(trips_by_person$wtr_trip)] <- 0
trips_by_person$dtr_trip[is.na(trips_by_person$dtr_trip)] <- 0
trips_by_person$active[is.na(trips_by_person$active)] <- 0

trips_by_person <- mutate(trips_by_person, more15=(active>=15), more30=(active>=30))

# summarize
active_summary <- summarise(group_by(trips_by_person, taz, county_name, ptype, zeroAuto), 
                            freq     = n(),
                            active   = mean(active),
                            more15   = mean(more15),
                            more30   = mean(more30),
                            wlk_trip = mean(wlk_trip),
                            bik_trip = mean(bik_trip),
                            wtr_trip = mean(wtr_trip),
                            dtr_trip = mean(dtr_trip),
                            atHomeA  = mean(atHomeA))

write.table(active_summary, file.path(TARGET_DIR,"core_summaries","ActiveTransport.csv"), sep=",", row.names=FALSE)
model_summary <- active_summary  # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR,"core_summaries","ActiveTransport.rdata"))
remove(active_summary, trips_by_person)
```

## Activity Pattern Summary

```{r ActivityPatternSummary, error=FALSE}
actpatt_summary <- summarise(group_by(persons, type, activity_pattern, imf_choice, inmf_choice, incQ_label), freq=n())
write.table(actpatt_summary, file.path(TARGET_DIR,"core_summaries","ActivityPattern.csv"), sep=",", row.names=FALSE)
model_summary <- actpatt_summary  # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR,"core_summaries","ActivityPattern.rdata"))
remove(actpatt_summary)
```

## Auto Ownership Summaries

```{r AutoOwnershipSummary, error=FALSE}
# summarise
autoown_summary <- summarise(group_by(households, SD, COUNTY, county_name, autos, 
                                      incQ, incQ_label, walk_subzone, walk_subzone_label, workers, kidsNoDr), freq=n())
write.table(autoown_summary, file.path(TARGET_DIR,"core_summaries","AutomobileOwnership.csv"), sep=",", row.names=FALSE)
model_summary <- autoown_summary  # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR,"core_summaries","AutomobileOwnership.rdata"))
```

### Auto Ownership Summary By County

```{r AutoOwnershipSummaryByCountyGraph, fig.width=8, fig.height=5}
auto_labels <- c("Zero automobiles", "One automobile", "Two automobiles",
                 "Three automobiles", "Four or more automobiles")

# by County
autoown_summary_county <- summarise(group_by(autoown_summary, COUNTY, county_name, autos), freq=sum(freq))
ggplot(autoown_summary_county, aes(x=as.factor(COUNTY), y=freq, fill=as.factor(autos))) +
  geom_bar(stat="identity", position="fill") +
  scale_y_continuous(labels=percent, breaks=seq(0,1,0.1)) +
  scale_x_discrete(labels=LOOKUP_COUNTY$county_name) +
  xlab("County of Residence") + 
  ylab("Share of Households") +
  ggtitle("Households by Automobile Ownership and County of Residence") +
  scale_fill_discrete(name="", 
                      breaks=c(4,3,2,1,0), 
                      labels=rev(auto_labels))
remove(autoown_summary_county)
```

### Auto Ownership Summary By Workers

```{r AutoOwnershipSummaryByWorkersGraph, fig.width=8, fig.height=5}
# by Workers
autoown_summary_workers <- summarise(group_by(autoown_summary, workers, autos), freq=sum(freq))

ggplot(autoown_summary_workers, aes(x=as.factor(workers), y=freq, fill=as.factor(autos))) +
  geom_bar(stat="identity", position="fill") +
  scale_y_continuous(labels=percent, breaks=seq(0,1,0.1)) +
  scale_x_discrete(labels=WORKER_LABELS) +
  xlab("Workers in the Household") + 
  ylab("Share of Households") +
  ggtitle("Households by Automobile Ownership and Number of Workers") +
  scale_fill_discrete(name="", 
                      breaks=c(4,3,2,1,0), 
                      labels=rev(auto_labels))
remove(autoown_summary, autoown_summary_workers)
```

## Commute By Employment Location Summaries
```{r CommuteByEmploymentLocationSummary, error=FALSE}
commute_tours   <- select(tours, hh_id, orig_taz, dest_taz, dest_COUNTY, dest_county_name, dest_SD,
                          tour_purpose, start_hour, end_hour, tour_mode, income, incQ, incQ_label, parking_cost)
# select out non-work travel
commute_tours   <- subset(commute_tours, (tour_purpose!="atwork_business" & tour_purpose!="atwork_eat"     &
                                          tour_purpose!="atwork_maint"    & tour_purpose!="eatout"         &
                                          tour_purpose!="escort_kids"     & tour_purpose!="escort_no kids" &
                                          tour_purpose!="othdiscr"        & tour_purpose!="othmaint"       &
                                          tour_purpose!="school_grade"    & tour_purpose!="school_high"    &
                                          tour_purpose!="shopping"        & tour_purpose!="social"         &
                                          tour_purpose!="university"))

# TODO: add this to all tours, not just commute?
# The cost database contains simplified transit modes, build a cross-walk from the full modes;
commute_tours   <- mutate(commute_tours, 
                          costMode  =(tour_mode<=8)*tour_mode +         # auto, walk, bike all in the skims
                                     # All walk to transit sub modes get best path walk to transit skim
                                     (tour_mode>=9)*(tour_mode<=13)*9 +
                                     # All drive transit walk sub modes get the best path
                                     # drive transit walk skim (tours start at home)
                                     (tour_mode>=14)*10,
                          timeCodeHwNum=(start_hour<6)*1 +                  # EA
                                        (start_hour>5)*(start_hour<10)*2 +  # AM
                                        (start_hour>9)*(start_hour<15)*3 +  # MD
                                        (start_hour>14)*(start_hour<19)*4 + # PM
                                        (start_hour>18)*5,                  # EV
                          timeCodeWhNum=(end_hour<6)*1 +                  # EA
                                        (end_hour>5)*(end_hour<10)*2 +    # AM
                                        (end_hour>9)*(end_hour<15)*3 +    # MD
                                        (end_hour>14)*(end_hour<19)*4 +   # PM
                                        (end_hour>18)*5                   # EV
                          )
# add timeCodeHw and timeCodeWh (character abbreviation versions)
commute_tours   <- left_join(commute_tours, 
                             select(mutate(LOOKUP_TIMEPERIOD, 
                                           timeCodeHwNum=timeCodeNum, timeCodeHw=timeperiod_abbrev), 
                                    timeCodeHwNum, timeCodeHw))
commute_tours   <- left_join(commute_tours, 
                             select(mutate(LOOKUP_TIMEPERIOD, 
                                           timeCodeWhNum=timeCodeNum, timeCodeWh=timeperiod_abbrev), 
                                    timeCodeWhNum, timeCodeWh))

# Initialize for home-work cost lookup
commute_tours   <- mutate(commute_tours, timeCode=timeCodeHw, cost=0.0, cost_fail=0)
# Add home-work cost to tours
for (timeperiod in LOOKUP_TIMEPERIOD$timeperiod_abbrev) {
  commute_tours <- add_cost(timeperiod, commute_tours, reverse_od=FALSE)
}
# That set it as cost, so rename
commute_tours   <- mutate(commute_tours, hwCost=cost)
# Add distance to tours
commute_tours   <- mutate(commute_tours, distance=0.0, distance_mode=tour_mode)
for (timeperiod in LOOKUP_TIMEPERIOD$timeperiod_abbrev) {
  commute_tours <- add_distance(timeperiod, commute_tours)
}

# Add time
commute_tours   <- mutate(commute_tours, time=0.0, time_fail=0)
for (timeperiod in LOOKUP_TIMEPERIOD$timeperiod_abbrev) {
  commute_tours <- add_time(timeperiod, commute_tours)
}


# Initialize for work-home cost lookup.
commute_tours   <- mutate(commute_tours, timeCode=timeCodeWh, cost=0.0)
# Add work-home cost to tours
for (timeperiod in LOOKUP_TIMEPERIOD$timeperiod_abbrev) {
  commute_tours <- add_cost(timeperiod, commute_tours, reverse_od=TRUE)
}
# That set it as cost, so rename
commute_tours   <- mutate(commute_tours, whCost=cost)
# Save cost as the average of hwCost and whCost
commute_tours   <- mutate(commute_tours, cost=0.5*(hwCost+whCost))

commute_tours   <- mutate(commute_tours, totCost=cost+parking_cost)

# tours with failures should be dropped from the mean
commute_tours$totCost[commute_tours$cost_fail>0] <- NA
commute_tours$cost[commute_tours$cost_fail>0]    <- NA
commute_tours$time[commute_tours$time_fail>0]    <- NA

# summarise
commute_summary <- summarise(group_by(commute_tours, dest_COUNTY, dest_county_name, dest_SD, tour_mode), 
                             freq         = n(),
                             totCost      = mean(totCost, na.rm=TRUE),
                             cost         = mean(cost,    na.rm=TRUE),
                             parking_cost = mean(parking_cost),
                             distance     = mean(distance),
                             cost_fail    = sum(cost_fail))
write.table(commute_summary, file.path(TARGET_DIR,"core_summaries","CommuteByEmploymentLocation.csv"),
            sep=",", row.names=FALSE)
model_summary <- commute_summary  # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR,"core_summaries","CommuteByEmploymentLocation.rdata"))
remove(commute_summary)
```

## Commute By Income Summaries
```{r CommuteByIncomeSummary, error=FALSE}
commute_inc_summary_byjob <- summarise(group_by(commute_tours, dest_COUNTY, dest_county_name, dest_SD, dest_taz, 
                                                tour_mode, incQ, incQ_label), 
                                       freq         = n(),
                                       totCost      = mean(totCost, na.rm=TRUE),
                                       cost         = mean(cost,    na.rm=TRUE),
                                       parking_cost = mean(parking_cost),
                                       distance     = mean(distance),
                                       duration     = mean(time,    na.rm=TRUE),
                                       cost_fail    = sum(cost_fail),
                                       time_fail    = sum(time_fail))
write.table(commute_inc_summary_byjob, 
            file.path(TARGET_DIR,"core_summaries","CommuteByIncomeJob.csv"), sep=",", row.names=FALSE)
model_summary <- commute_inc_summary_byjob  # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR,"core_summaries","CommuteByIncomeJob.rdata"))

# The COUNTY and super district are by destination right now; add residence
commute_tours   <- left_join(commute_tours,
                             mutate(tazData, orig_taz=taz, res_COUNTY=COUNTY, res_county_name=county_name, 
                                    res_SD=SD) %.% 
                               select(orig_taz, res_COUNTY, res_county_name, res_SD),
                             by=c("orig_taz"))
commute_inc_summary_byres <- summarise(group_by(commute_tours, res_COUNTY, res_county_name, res_SD, orig_taz,
                                                tour_mode, incQ, incQ_label), 
                                       freq         = n(),
                                       totCost      = mean(totCost, na.rm=TRUE),
                                       cost         = mean(cost,    na.rm=TRUE),
                                       parking_cost = mean(parking_cost),
                                       distance     = mean(distance),
                                       duration     = mean(time,    na.rm=TRUE),
                                       cost_fail    = sum(cost_fail),
                                       time_fail    = sum(time_fail))
write.table(commute_inc_summary_byres, 
            file.path(TARGET_DIR,"core_summaries","CommuteByIncomeHousehold.csv"), sep=",", row.names=FALSE)
model_summary <- commute_inc_summary_byres  # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR,"core_summaries","CommuteByIncomeHousehold.rdata"))
remove(commute_tours, commute_inc_summary_byjob, commute_inc_summary_byres)
```

## Journey To Work Summary
```{r JourneyToWorkSummary, error=FALSE}
# select out non-work travel
work_locations <- subset(mandatory_locations, WorkLocation != 0)
# add home county
work_locations <- left_join(work_locations, 
                            mutate(tazData,HomeTAZ=taz, homeCOUNTY=COUNTY, home_county_name=county_name) %.% 
                              select(HomeTAZ,homeCOUNTY,home_county_name),
                            by=c("HomeTAZ"))
# add work county
work_locations <- left_join(work_locations, 
                            mutate(tazData,WorkLocation=taz, workCOUNTY=COUNTY, work_county_name=county_name) %.% 
                              select(WorkLocation,workCOUNTY,work_county_name),
                            by=c("WorkLocation"))

journeytowork_summary <- summarise(group_by(work_locations,
                                            homeCOUNTY, home_county_name, HomeTAZ,
                                            WorkLocation, workCOUNTY, work_county_name),
                                   freq   = n(),
                                   Income = mean(Income))
write.table(journeytowork_summary,
            file.path(TARGET_DIR,"core_summaries","JourneyToWork.csv"), sep=",", row.names=FALSE)
model_summary <- journeytowork_summary  # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR,"core_summaries","JourneyToWork.rdata"))
remove(mandatory_locations, work_locations, journeytowork_summary, model_summary)
```

## Time of Day Summary
```{r TimeOfDaySummary, error=FALSE}
timeofday_summary <- summarise(group_by(tours,SD,COUNTY,county_name,
                                        simple_purpose,tour_mode,start_hour,end_hour),
                               freq              = n(),
                               num_participants  = sum(num_participants))
write.table(timeofday_summary,
            file.path(TARGET_DIR,"core_summaries","TimeOfDay.csv"), sep=",", row.names=FALSE)
model_summary <- timeofday_summary # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR, "core_summaries", "TimeOfDay.rdata"))

# another table -- who's touring?
persons_touring <- data.frame(hour=numeric(), simple_purpose=character(), persons_touring=numeric())
for (h in min(timeofday_summary$start_hour):max(timeofday_summary$end_hour)) { 
  touring_at_hour <- subset(timeofday_summary, 
                         (timeofday_summary$start_hour<=h)&(timeofday_summary$end_hour>=h))
  touring_at_hour <- summarise(group_by(touring_at_hour, simple_purpose), persons_touring=sum(num_participants))
  touring_at_hour <- mutate(touring_at_hour, hour=h)
  persons_touring <- rbind(persons_touring, touring_at_hour)
}
write.table(persons_touring,
            file.path(TARGET_DIR,"core_summaries","TimeOfDay_personsTouring.csv"), 
            sep=",", row.names=FALSE)
model_summary <- persons_touring
save(model_summary, file=file.path(TARGET_DIR, "core_summaries", "TimeOfDay_personsTouring.rdata"))
remove(timeofday_summary, persons_touring, h, touring_at_hour)
```

## Trip Distance Summary
```{r TripDistanceSummary, error=FALSE}
tripdist_summary          <- summarise(group_by(trips, autoSuff, autoSuff_label, incQ, incQ_label, timeCode,
                                                timeperiod_label, trip_mode, tour_purpose),
                                       freq         = n(),
                                       distance     = mean(distance))
write.table(tripdist_summary,
            file.path(TARGET_DIR,"core_summaries","TripDistance.csv"), sep=",", row.names=FALSE)
model_summary <- tripdist_summary # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR, "core_summaries", "TripDistance.rdata"))
remove(tripdist_summary, model_summary)
```

## Cleanup and save persons
```{r savePersons, error=FALSE}
save(persons, file=file.path(TARGET_DIR, "updated_output", "persons.rdata"))
remove(persons)
```

## Travel Cost Summary
```{r TravelCostSummary, error=FALSE}

# split trip cost into trip_cost_indiv, trip_cost_joint
trips       <- mutate(trips,
                      trip_cost_indiv = (num_participants==1)*cost,
                      trip_cost_joint = (num_participants>1)*cost/num_participants
                      )

# we only want some variables - sum them to hh_id
trips_melt  <- tbl_df(melt(select(trips, hh_id, trip_cost_indiv, trip_cost_joint, cost_fail), 
                           id.vars=c("hh_id")))
trips_dcast <- tbl_df(dcast(trips_melt, hh_id ~ variable, fun.aggregate=sum))
# left join with household --> one row per household!
costs_by_household <- left_join(select(households, hh_id, SD, COUNTY, county_name, 
                                       PERSONS, incQ, incQ_label, autos), 
                             trips_dcast, by=c("hh_id"))

# we're done with these
remove(trips_melt,trips_dcast)

# recode people = PERSONS capped at 6
costs_by_household <- mutate(costs_by_household, people=ifelse(PERSONS>6,6,PERSONS))

# fill in NA
costs_by_household$trip_cost_indiv[is.na(costs_by_household$trip_cost_indiv)] <- 0
costs_by_household$trip_cost_joint[is.na(costs_by_household$trip_cost_joint)] <- 0
costs_by_household$cost_fail[is.na(costs_by_household$cost_fail)] <- 0

# parking cost is in tours
tours       <- mutate(tours,
                      pcost_indiv = (num_participants==1)*parking_cost,
                      pcost_joint = (num_participants>1)*parking_cost)
# sum them to hh_id
tours_melt  <- tbl_df(melt(select(tours, hh_id, pcost_indiv, pcost_joint), 
                           id.vars=c("hh_id")))
tours_dcast <- tbl_df(dcast(tours_melt, hh_id ~ variable, fun.aggregate=sum))

# add to costs_by_household
costs_by_household <- left_join(costs_by_household, tours_dcast, by=c("hh_id"))

# fill in NA
costs_by_household$pcost_indiv[is.na(costs_by_household$pcost_indiv)] <- 0
costs_by_household$pcost_joint[is.na(costs_by_household$pcost_joint)] <- 0

costs_by_household <- mutate(costs_by_household,
                             total_cost=trip_cost_indiv+trip_cost_joint+
                                        pcost_indiv+pcost_joint)
remove(tours_melt, tours_dcast)

# write summary
travelcost_summary  <- summarise(group_by(costs_by_household, SD, COUNTY, county_name,
                                          people, incQ, incQ_label, autos),
                                 freq            = n(),
                                 total_cost      = mean(total_cost),
                                 trip_cost_indiv = mean(trip_cost_indiv),
                                 trip_cost_joint = mean(trip_cost_joint),
                                 cost_fail       = sum(cost_fail),
                                 pcost_indiv     = mean(pcost_indiv),
                                 pcost_joint     = mean(pcost_joint))
write.table(travelcost_summary,
            file.path(TARGET_DIR,"core_summaries","TravelCost.csv"), sep=",", row.names=FALSE)
model_summary <- travelcost_summary # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR, "core_summaries", "TravelCost.rdata"))
remove(costs_by_household,travelcost_summary,model_summary)
```

## Per Trip Travel Time Summary
```{r PerTripTravelTimeSummary, error=FALSE}
triptime_summary <- summarise(group_by(trips, incQ, incQ_label, trip_mode, tour_purpose),
                              freq             = n(),
                              num_participants = sum(num_participants),
                              trvlTime         = mean(time),
                              time_fail        = sum(time_fail))
write.table(triptime_summary,
            file.path(TARGET_DIR,"core_summaries","PerTripTravelTime.csv"), sep=",", row.names=FALSE)
model_summary <- triptime_summary # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR, "core_summaries", "PerTripTravelTime.rdata"))
remove(triptime_summary, model_summary)
```

## Cleanup and save tours, trips and households
```{r saveToursAndTrips, error=FALSE}
save(trips, file=file.path(TARGET_DIR, "updated_output", "trips.rdata"))
remove(trips)

save(tours, file=file.path(TARGET_DIR, "updated_output", "tours.rdata"))
remove(tours)

save(households, file=file.path(TARGET_DIR, "updated_output", "households.rdata"))
remove(households)
```

## Vehicle Miles Traveled
```{r VehicleMilesTraveledSummary, error=FALSE}
# maybe don't save/load but just keep it?  I started having memory problems though...
load(file=file.path(TARGET_DIR,"updated_output","trips.rdata"))
auto_trips <- subset(trips, trip_mode<=6)
remove(trips)

auto_trips <- mutate(auto_trips,
                     vmt_indiv=(num_participants==1)*(trip_mode==1)*distance +
                               (num_participants==1)*(trip_mode==2)*distance +
                               (num_participants==1)*(trip_mode==3)*(distance/2.0) +
                               (num_participants==1)*(trip_mode==4)*(distance/2.0) +
                               (num_participants==1)*(trip_mode==5)*(distance/3.25) +
                               (num_participants==1)*(trip_mode==6)*(distance/3.25),
                     vmt_joint=(num_participants>1)*(distance/num_participants),
                     vmt=vmt_indiv+vmt_joint)

# sum them to hh_id, person_id
auto_trips_melt  <- tbl_df(melt(select(auto_trips, hh_id, person_id, vmt_indiv, vmt_joint, vmt), 
                           id.vars=c("hh_id", "person_id")))
auto_trips_dcast <- tbl_df(dcast(auto_trips_melt, hh_id+person_id ~ variable, fun.aggregate=sum))
remove(auto_trips_melt)

# add to persons so we get zero-vmt persons
load(file=file.path(TARGET_DIR,"updated_output","persons.rdata"))
persons          <- left_join(persons, auto_trips_dcast, by=c("hh_id","person_id"))
remove(auto_trips_dcast)
persons$vmt_indiv[is.na(persons$vmt_indiv)] <- 0
persons$vmt_joint[is.na(persons$vmt_joint)] <- 0
persons$vmt[is.na(persons$vmt)]             <- 0

# add household vars
load(file=file.path(TARGET_DIR,"updated_output","households.rdata"))
persons          <- left_join(persons, select(households, hh_id, taz, COUNTY, county_name,
                                              walk_subzone, walk_subzone_label, 
                                              autoSuff, autoSuff_label),
                              by=c("hh_id"))
remove(households)

# added new person vars so save it
save(persons, file=file.path(TARGET_DIR, "updated_output", "persons.rdata"))

vmt_summary      <- summarise(group_by(persons, COUNTY, county_name, taz,
                                       walk_subzone, walk_subzone_label,
                                       ptype, ptype_label,
                                       autoSuff, autoSuff_label),
                              freq             = n(),
                              vmt_indiv        = mean(vmt_indiv),
                              vmt_joint        = mean(vmt_joint),
                              vmt              = mean(vmt))
write.table(vmt_summary,
            file.path(TARGET_DIR,"core_summaries","VehicleMilesTraveled.csv"), sep=",", row.names=FALSE)
model_summary <- vmt_summary # name it generically for rdata
save(model_summary, file=file.path(TARGET_DIR, "core_summaries", "VehicleMilesTraveled.rdata"))
remove(vmt_summary, model_summary)
```