---
title: "Build Database"
author: "David Ory"
output: 
  html_document:
    theme: cosmo
    toc: yes
---


## Administration

#### Purpose
This script consumes Cube Voyager TRNBUILD link assignments and extracts BART boardings by BART station and access mode (walk or drive).  These are summaries from MTC's Travel Model One. 

#### Outputs
1.  A CSV database of BART boardings by BART station, travel model scenario, access mode, time of day, and model path.  

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
library(foreign)
suppressMessages(library(dplyr))
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote I/O locations
```{r remote-io}
D_INPUT  = "M:/Application/Model One/Value Parking Pricing Pilot/Scenarios/"
D_OUTPUT = "M:/Application/Model One/Value Parking Pricing Pilot/Across Scenarios/"
F_OUTPUT = paste(D_OUTPUT, "bart_boardings_by_access_mode.csv", sep = "") 

F_SCENARIO_NAMES = "M:/Application/Model One/Value Parking Pricing Pilot/Scenarios/ScenarioKey.csv" 

SCENARIO_ID_LIST = c("2015_04_137", "2015_04_138", "2015_04_139", "2015_04_140", "2015_04_141", "2015_04_142")
PATH_DBF_LIST    = c("wlk_hvy_wlk.dbf", "drv_hvy_wlk.dbf", "wlk_hvy_drv.dbf",
                     "wlk_com_wlk.dbf", "drv_com_wlk.dbf", "wlk_com_drv.dbf")
TIME_PERIOD_LIST = c("EA", "AM", "MD", "PM", "EV")

# BART station nodes file
F_BART_STATION_NODES = "bart-station-nodes.csv"

```

#### Function to extract needed data as DBFs are read into memory
```{r filter-function}
Filter_Data <- function(input_df){
  
  output_df <- input_df %>%
    select(A, B, MODE, AB_VOL) %>%
    filter(MODE == 3 | MODE == 4 | MODE == 5) %>%
    filter(AB_VOL > 0)
  
  output_df <- left_join(output_df, bart_nodes, by = c("B"))
  
  output_df <- output_df %>%
    filter(!is.na(Station))
  
  return(output_df)
  
}

```

#### Data read and filter
```{r data-read-filter}
scenario_names <- read.table(file = F_SCENARIO_NAMES, header = TRUE, sep = ",", stringsAsFactors = FALSE)
bart_nodes <- read.table(file = F_BART_STATION_NODES, header = TRUE, sep = ",", stringsAsFactors = FALSE)

bart_nodes <- bart_nodes %>%
  select(Station, B = Node)

for (scenario_index in 1:length(SCENARIO_ID_LIST)) {
  for (path_index in 1:length(PATH_DBF_LIST)) {
    for (time_index in 1:length(TIME_PERIOD_LIST)) {
      
      input_file_name <- paste(D_INPUT, 
                               SCENARIO_ID_LIST[scenario_index], 
                               "/OUTPUT/trn/trnlink",  
                               TIME_PERIOD_LIST[time_index], 
                               "_",
                               PATH_DBF_LIST[path_index], sep = "")
      
      input <- read.dbf(file = input_file_name, as.is = TRUE)
      
      filtered <- Filter_Data(input)
      
      filtered <- filtered %>%
        mutate(path        = substr(PATH_DBF_LIST[path_index], 1, 11)) %>%
        mutate(time_period = TIME_PERIOD_LIST[time_index]) %>%
        mutate(src         = SCENARIO_ID_LIST[scenario_index])
      
      if (scenario_index == 1 & path_index == 1 & time_index == 1) {
        data <- filtered
        }
      else {
        data <- rbind(data, filtered)
        }
      
      } # time period
    } # path
  } # scenario

```

#### Join the scenario names
```{r joins}
data <- left_join(data, scenario_names, by = c("src"))
```

#### Write to disk
```{r write}
write.csv(data, file = F_OUTPUT, row.names = FALSE, quote = F)
```


