---
title: "Check Zonal Data"
author: "David Ory"
runtime: shiny
output: 
   html_document:
      theme: cosmo
      toc: yes
---

## Administration

#### Purpose
This script consumes TAZ data created by UrbanSim or ABAG and processed by the Java program MtcProcessAbagData.  The script adds a year to each file and spits out a single database for use in Tableau -- see check-zonal-data.twb in this directory.  

#### Outputs
1.  consolidate database

## Procedure

#### Overhead
```{r overhead, results = 'hide', include=FALSE}
library(knitr)
suppressMessages(library(dplyr))
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Remote I/O
```{r remote-io, include=FALSE}
F_WORKING_DIR = "M:/Application/Model One/Zonal Data/Plan Bay Area 2040/2015 11 23 UrbanSim Tests/"

F_OUTPUT = paste(F_WORKING_DIR, "consolidated-taz.csv", sep = "") 
```

#### Data read method
```{r data-read-method}
Read_Data <- function(string_file_location, integer_year) {
  
  working_df <- read.table(file = string_file_location, header = TRUE, sep = ",", stringsAsFactors = FALSE)
  
  working_df <- working_df %>%
    mutate(year = integer_year)
  
  return(working_df)
}

```

#### Data reads and concatenate
```{r data-reads}
working <- Read_Data(paste(F_WORKING_DIR, "tazData2010.csv", sep = ""), 2010)
running <- working

working <- Read_Data(paste(F_WORKING_DIR, "tazData2015.csv", sep = ""), 2015)
running <- rbind(running, working)

working <- Read_Data(paste(F_WORKING_DIR, "tazData2020.csv", sep = ""), 2020)
running <- rbind(running, working)

working <- Read_Data(paste(F_WORKING_DIR, "tazData2025.csv", sep = ""), 2025)
running <- rbind(running, working)

working <- Read_Data(paste(F_WORKING_DIR, "tazData2030.csv", sep = ""), 2030)
running <- rbind(running, working)

working <- Read_Data(paste(F_WORKING_DIR, "tazData2035.csv", sep = ""), 2035)
running <- rbind(running, working)

working <- Read_Data(paste(F_WORKING_DIR, "tazData2040.csv", sep = ""), 2040)
running <- rbind(running, working)

```


#### Data write
```{r data-write}
write.csv(running, file = F_OUTPUT, row.names = FALSE, quote = F)

```

