---
title: "Longitudinal Observed and Estimated VMT Per Worker"
author: "David Ory"
output: 
   html_document:
      theme: cosmo
      toc: yes
---

## Administration

#### Purpose
This script consumes observed VMT per worker by workplace travel analysis zone estimates from the 2000 and 2012/13 household travel surveys and simulated VMT per worker by workplace estimates from the travel model.  These data is put into a single output database that is handled in Tableau. 


#### Outputs
1.  Seems to be working

#### _TODO_ 
1. Bug checks 
3. Add read me

## Procedure

#### Overhead
```{r overhead, results = 'hide'}
library(knitr)
library(SDMTools)
suppressMessages(library(dplyr))
```

```{r config, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

#### Parameters
```{r parameters}
YEAR_2000_DETAIL = "2000_05_002"
YEAR_2005_DETAIL = "2005_05_003"
YEAR_2010_DETAIL = "2010_05_003"

OBSERVED_CORRELATION_THRESHOLD = 30

```


#### Remote data locations
```{r remote-io}
F_OBSERVED_2000 = "M:/Data/HomeInterview/Longitudinal VMT per Capita and per Worker/BATS2000_VMT_by_Person.csv"
F_OBSERVED_2013 = "M:/Data/HomeInterview/Longitudinal VMT per Capita and per Worker/CHTS2013_VMT_by_Person.csv"

F_ESTIMATED_2000 = "M:/Application/Model One/RTP2017/Scenarios/2000_05_002/OUTPUT/core_summaries/AutoTripsVMT_perOrigDestHomeWork.csv"
F_ESTIMATED_2005 = "M:/Application/Model One/RTP2017/Scenarios/2005_05_003/OUTPUT/core_summaries/AutoTripsVMT_perOrigDestHomeWork.csv"
F_ESTIMATED_2010 = "M:/Application/Model One/RTP2017/Scenarios/2010_05_003/OUTPUT/core_summaries/AutoTripsVMT_perOrigDestHomeWork.csv"

F_EST_PERSONS_2000 = "M:/Application/Model One/RTP2017/Scenarios/2000_05_002/OUTPUT/core_summaries/AutoTripsVMT_personsHomeWork.csv"
F_EST_PERSONS_2005 = "M:/Application/Model One/RTP2017/Scenarios/2005_05_003/OUTPUT/core_summaries/AutoTripsVMT_personsHomeWork.csv"
F_EST_PERSONS_2010 = "M:/Application/Model One/RTP2017/Scenarios/2010_05_003/OUTPUT/core_summaries/AutoTripsVMT_personsHomeWork.csv"

F_TAZ_XY_COORDS = "~/GitHub/travel-model-one/utilities/taz-xy-coordinates/taz_nodes_WGS84.csv"

F_TAZ_COUNTY_CROSSWALK = "~/GitHub/travel-model-one/utilities/geographies/taz-superdistrict-county.csv"

F_OUTPUT = "~/../Box Sync/Share Data/observed-estimated-vmt/long-obs-est-vmt-per-worker.csv"

```

#### Data reads
```{r data-reads}
obs_2000 <- read.table(file = F_OBSERVED_2000, header = TRUE, sep = ",", stringsAsFactors = FALSE)
obs_2013 <- read.table(file = F_OBSERVED_2013, header = TRUE, sep = ",", stringsAsFactors = FALSE)

est_2000 <- read.table(file = F_ESTIMATED_2000, header = TRUE, sep = ",", stringsAsFactors = FALSE)
est_2005 <- read.table(file = F_ESTIMATED_2005, header = TRUE, sep = ",", stringsAsFactors = FALSE)
est_2010 <- read.table(file = F_ESTIMATED_2010, header = TRUE, sep = ",", stringsAsFactors = FALSE)

est_pers_2000 <- read.table(file = F_EST_PERSONS_2000, header = TRUE, sep = ",", stringsAsFactors = FALSE)
est_pers_2005 <- read.table(file = F_EST_PERSONS_2005, header = TRUE, sep = ",", stringsAsFactors = FALSE)
est_pers_2010 <- read.table(file = F_EST_PERSONS_2010, header = TRUE, sep = ",", stringsAsFactors = FALSE)

taz_xy <- read.table(file = F_TAZ_XY_COORDS, header = TRUE, sep = ",", stringsAsFactors = FALSE)

taz_sd_county <- read.table(file = F_TAZ_COUNTY_CROSSWALK, header = TRUE, sep = ",", stringsAsFactors = FALSE)

```

#### Manipulate taz xy
```{r manip-taz}
taz_df <- taz_xy %>%
  select(taz = N, latitude, longitude)

taz_sd_county_df <- taz_sd_county %>%
  select(taz = ZONE, super_district = SD_NAME, county = COUNTY_NAME)
```


#### Manipulate observed
```{r manip-obs}
working_2000 <- obs_2000 %>%
  mutate(year = 2000) %>%
  mutate(detail = "BATS")

working_2013 <- obs_2013 %>%
  mutate(year = 2013) %>%
  mutate(detail = "CHTS")

working <- rbind(working_2000, working_2013)

working_observed <- working %>%
  select(year, 
         detail, 
         taz = worktaz, 
         estimated_persons = hhweight, 
         vmt = weighted_vmt,
         id = hhid) %>%
  mutate(observed_persons = 1.0) %>%
  filter(!is.na(taz))

working_observed_sum <- working_observed %>%
  group_by(year, detail, taz) %>%
  summarise(mean_vmt = wt.mean(vmt, estimated_persons),
            sd_vmt   = wt.sd(vmt, estimated_persons),
            sum_vmt  = sum(vmt),
            sum_estimated_persons = sum(estimated_persons),
            sum_observed_persons  = sum(observed_persons)) %>%
  select(year, detail, taz, observed_persons = sum_observed_persons, estimated_persons = sum_estimated_persons, mean_vmt, sd_vmt, sum_vmt) %>%
  mutate(vmt_per_worker = sum_vmt / estimated_persons)

remove(working_2000, working_2013, working, working_observed)

```

#### Manipulate estimated
```{r manip-est}
# get persons by workplace location
working_2000 <- est_pers_2000 %>%
  mutate(year = 2000) %>%
  mutate(detail = YEAR_2000_DETAIL)

working_2005 <- est_pers_2005 %>%
  mutate(year = 2005) %>%
  mutate(detail = YEAR_2005_DETAIL)

working_2010 <- est_pers_2010 %>%
  mutate(year = 2010) %>%
  mutate(detail = YEAR_2010_DETAIL)

working <- rbind(working_2000, working_2005, working_2010)

working_pers_sum <- working %>%
  select(year, detail, taz = WorkLocation, freq) %>%
  filter(!(taz == 0)) %>%
  group_by(year, detail, taz) %>%
  summarise(estimated_persons = sum(freq))


# get VMT data
working_2000 <- est_2000 %>%
  mutate(year = 2000) %>%
  mutate(detail = YEAR_2000_DETAIL)

working_2005 <- est_2005 %>%
  mutate(year = 2005) %>%
  mutate(detail = YEAR_2005_DETAIL)

working_2010 <- est_2010 %>%
  mutate(year = 2010) %>%
  mutate(detail = YEAR_2010_DETAIL)

working <- rbind(working_2000, working_2005, working_2010)

working_vmt_sum <- working %>%
  select(year, detail, taz = WorkLocation, vmt) %>%
  filter(!(taz == 0)) %>%
  group_by(year, detail, taz) %>%
  summarise(sum_vmt = sum(vmt))

working_estimated_sum <- left_join(working_pers_sum, working_vmt_sum, by = c("year", "detail", "taz"))
  
working_estimated_sum <- working_estimated_sum %>%
  mutate(observed_persons = NA) %>%
  mutate(mean_vmt = NA) %>%
  mutate(sd_vmt = NA) %>%
  mutate(vmt_per_worker = ifelse(estimated_persons > 0L, sum_vmt / estimated_persons, NA))


remove(working_2000, working_2005, working_2010, working, working_pers_sum, working_vmt_sum)
  

```

#### Combine & Analyze
```{r combine}
outcome_df <- rbind(working_observed_sum, working_estimated_sum)

# Check correlations
# Year 2000
observed <- outcome_df %>%
  ungroup() %>%
  filter(detail == "BATS" & observed_persons > OBSERVED_CORRELATION_THRESHOLD) %>%
  select(taz, observed_vmt_per_worker = vmt_per_worker)

estimated <- outcome_df %>%
  ungroup() %>%
  filter(detail == YEAR_2000_DETAIL) %>%
  select(taz, estimated_vmt_per_worker = vmt_per_worker)

correlation <- left_join(observed, estimated, by = c("taz"))

cor(correlation$observed_vmt_per_worker, correlation$estimated_vmt_per_worker)

# Year 2010ish
observed <- outcome_df %>%
  ungroup() %>%
  filter(detail == "CHTS" & observed_persons > OBSERVED_CORRELATION_THRESHOLD) %>%
  select(taz, observed_vmt_per_worker = vmt_per_worker)

estimated <- outcome_df %>%
  ungroup() %>%
  filter(detail == YEAR_2010_DETAIL) %>%
  select(taz, estimated_vmt_per_worker = vmt_per_worker)

correlation <- left_join(observed, estimated, by = c("taz"))

cor(correlation$observed_vmt_per_worker, correlation$estimated_vmt_per_worker)


```

#### Write to disk
```{r data-writes}
outcome_df <- left_join(outcome_df, taz_df, by = c("taz"))
outcome_df <- left_join(outcome_df, taz_sd_county_df, by = c("taz"))

write.csv(outcome_df, file = F_OUTPUT, row.names = FALSE, quote = F)

```